---
title: "Closing the Gender Gap in Climate"
subtitle: "Can Electing More Women into Office Narrow the Climate Divide?"
author: "Sam Lee"
format: pdf
header-includes:
  - \usepackage{fontspec}
  - \setmainfont{Times New Roman}
  - \usepackage{setspace}
  - \renewcommand{\normalsize}{\fontsize{12}{14.4}\selectfont}
  - \linespread{2}
  - \usepackage[margin=1in]{geometry}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)

library(tidyverse)
library(patchwork)
library(stargazer)
library(haven)
library(ggrepel)

women <- read_dta("Data/womenrep.dta")
```

# Abstract

I seek to identify the causal effect that electing higher proportions of women in national legislatures in African and Arab nations has on yearly per capita $CO_2$ emissions. I utilize a two-stage difference-in-difference approach on 64 different countries throughout Africa and the Middle-East from 1998-2022. To eliminate the potential endogeneity in using the proportion of women in national legislature as a treatment on per capita $CO_2$ emissions, I use years since women were granted suffrage in each respective country as an instrumental variable. One nominal paper in the literature has shown evidence that increased proportions of women in parliaments is more than likely causally related to stricter climate policies (Mavisakalyan & Tarverdi, 2019). However, this paper does not take into account how carbon emissions and women in government have evolved over time. To my knowledge, little research has been done to show a causal link between women in parliaments and its effects on climate *over time*. I expand on this research by showing how carbon emissions have changed over time to elicit a local average treatment effect between changes in gender compositions at the national legislature level. When the average proportion of women in national legislatures in African and Arab countries increase by 1%, I find a significant effect that total yearly (gigatons of) $CO_2$ emissions per capita decrease by 1.08 (-1.64, -0.51), on average. To enforce the credibility of significance, I run a placebo regression, using the number of natural disasters (as classified and recorded by the Centre for Research on the Epidemiology of Disasters) that occur within a country at a particular year as an effective placebo.

# Introduction

The current climate and economic research today overwhelmingly supports the thesis that women---especially in developing countries---are disproportionately affected by the negative externalities of climate change. During periods of increased droughts, which are exacerbated due to the effects of climate change, women often make long trips to meet agricultural, hygiene, and family needs. This has been linked to increased violence against women (UNDP, 2019). Increased risks of poverty, food insecurities have also been shown to increase as a result of increased rates of natural disasters and variable weather patterns. One article reports that, "when a family is faced with the impact of the climate crisis, girls’ education is one of the first things families drop" (Medlicott, 2021). Another meta-analysis consisting of 53 studies report that two-thirds of those studies find that women are more often the victims of death or injury in the case of extreme weather events (Seller, 2016). If women are, on average, affected disproportionately affected by climate externalities, then one would expect that, on average, rational policymakers who are women would enact effective climate policy at higher rates (pertaining to protecting the needs of women) relative to other compositions of policymakers with lower proportions of women, assuming policymakers who are women have the means and power to do so.

The outline of this paper will proceed as follows: I will first introduce the data I collected for this study and assess its reliability. I will then discuss the identification strategy I use to to estimate the local average treatment effect. I support this result with a placebo analysis, and then I will finish by discussing the results of my econometric analysis, its implications, and then conclude.

# Data

The data used for this analysis was strategically gathered[^1] using publicly available sources (either by ethical web-scraping methods or direct download). The Inter-Parliamentary Union (IPU) maintains a consistent archive of records that monitors the number of women elected into the national legislature for each country. Not every government maintains the same structure of government, however, the IPU distinguishes between women elected into the lower-house and upper-house levels of government throughout multiple time periods. I take the average proportion of women holding office in the national legislature overall in a given year by taking a weighted average of the two houses[^2][^3]. I obtained emissions data, including total yearly (gigatons of) $CO_2$ emissions per capita per country, from the Emissions Database for Global Atmospheric Research (EDGAR) as maintained by the European Commission[^4]. I obtained country-year-level specific covariates, such as population and GDP, from the World Bank[^5]. I obtained the data base linking each country's year they passed a right of universal suffrage from "A Lexical Index of Electoral Democracy" on the Harvard Dataverse (Skaaning et. al, 2015). I checked country-specific anomalies using research from the Pew Research Center[^6] (Schaeffer, 2020). Finally, I used U.S. Department of State's regional definitions to define which countries I define as either African or Arab.

[^1]: Further documentation and the scripts used to obtain the data used for this analysis can be found on this project's Github repository [https://github.com/SamLeeBYU/Elections](https://github.com/SamLeeBYU/Elections).
[^2]: Note that econometrically, this assumes that on average, the power that women are given in each house of national legislature is equal to the 'weight' of women in that chamber of legislature: In other words, assuming a fair political system, I assume no chamber can systemically dominate the other chamber.
[^3]: A graphical illustration of the average proportion of women in the national legislatures for each country in my sample can be seen in Figure 1 of the Appendix.
[^4]: A graphical illustration of the yearly $CO_2$ emissions over time for each country in my sample can be seen in Figure 2 of the Appendix.
[^5]: Statistics for these key variables are summarized in Table 1 of the Appendix.
[^6]: Not all countries in Africa and the Middle-East treat women equally. Even when suffrage is granted, it is not granted universally. Thus, the instrument may not influence the proportion of women in government the same in every country. This does not discredit my econometric analysis, however, so long as the monotonicity assumption of my instrument is met (which will be discussed hereafter). However, I exclude obvious anomalies in my analytic sample. See Table 2 in the Appendix for summary statistics for each country in my sample.

By adjusting for time, country, and other specific endogenous effects on the proportion of women in government in a country ($c$) at year ($t$), I hope to show that effectively, when proportions of women in government increase, $CO_2$ emissions decrease. However, one potential problem with data is that I do not account for country-specific government intricacies that could prevent prevent women or a legislature from passing effective climate policy---such as not accounting for government corruption or other unique anomalies (wars between countries, coups, laws that are discriminatory against women, etc, that prevent a government from acting efficiently)---at every year[^7].

[^7]: To assert consistency in my estimates, econometrically, I assume that on average, these factors in the idiosyncratic term of my econometric model are 0 for any given country at year $t$. (See equation (3)).

In my sample, the United Arab Emirates (UAE) and Saudia Arabia are excluded because they don't give full suffrage to women: The UAE held its first national elections in 2006, but only a select amount of men and women were eligible to participate (Schaeffer, 2020). Saudi Arabia doesn't hold national elections[^8]. This leaves us with a total of 64 different countries in the analytic sample. Merging the suffrage data with all the IPU available data and the $CO_2$ emissions data from EDGAR, I was left with data spanning 1997-2022[^9].

[^8]: Women are, however, enfranchised on the local level government.
[^9]: Notably, the data is left as an unbalanced panel data set (See Table 2 in the Appendix). The World Bank does not have GDP or population data for every single year of every single country in the data, and critically, the IPU also notes that it cannot reliably retrieve the number of women in national legislatures for every year. These years for those specific countries are naturally excluded. I assume that the probability at which data for a specific year-country combination will be missing is essentially random. This assumption may be violated if countries that are systemically difficult to obtain data for (such as Somalia or the Central African Republic) is also correlated with other covariates and or $CO_2$ emissions.

# Identification

To estimate the causal effect that electing more women in national legislatures in African and Arab countries has on reducing per capita $CO_2$ emissions, I employ a two-stage difference-in-difference identification strategy. In regards to curbing the negative externalities of climate change in the context of this paper, I use per capita $CO_2$ emissions as a response variable. Let $Y_{ct}$ represent the total (gigatons of) carbon emissions per capita for a country ($c$) in a year ($t$). I seek to identify consistently and interpret the local average treatment effect of $\delta$, the effect of higher proportion of women in national legislatures in African and Arab countries on national $CO_2$ emissions per capita. Research from Mirziyoyeva and Salahodjaev (2023) suggests that women in government promote economic growth. This could pose potential problems of reverse causality: We need to be certain that any effect we see on women representation causes a reduction (or increase) in carbon emissions. If carbon emissions are highly correlated with economic growth and industrialization then an OLS regression (that is, using a typical difference-in-differences approach) on $Y_{ct}$ to estimate $\delta$ would almost certainly be biased and inconsistent. To first eliminate the endogeneity for reverse causality---the supposition that perhaps lower carbon emissions causes a higher proportion of women to become elected during the same year; or rather, perhaps higher carbon emissions are correlated with other variables such as economic growth that are linked to higher proportions of women becoming elected---I lag the treatment variable of interest, $W_{ct}$, the average proportion of women in national legislature for country $c$ during year $t$ one year. This implies that any change in $Y_{ct}$ will inherently come *after* a change in $W_{ct}$. While I do not focus on specific regulatory practices within this paper, effective climate policy would consist of enacting particular policies and regulations that would reduce the $CO_2$ emissions by economic constraint. Hence, as a matter of the causal chain of inference, it could be that increased proportions of women in government could (causally) affect economic performance within a country at time $t$. Thus, to establish the causal chain of inference, I infer:
$$
\sum_{p=2}^PX_{ct-p}\rightarrow W_{ct-1}\rightarrow Y_{ct}
$$
Where $X_{ct-p}$ is a matrix of controls related to economic growth. For the same reason the $W_{ct}$ is lagged one year to ensure that a change in $Y_{ct}$ comes after a change in $W_{ct}$, I include a series of autoregressive lags in the economic controls (starting at lag 2) to ensure that changes in $W_{ct-1}$ occur *after* changes in economic development. A priori, I set $P=5$ in my econometric model to control for economic lags in a country. For the purpose of this two-stage regression, $X_{ct-p}$ consists of logged GDP and logged population for a country at year $t$[^10].

[^10]: A further analysis could (and should) include further economic controls that may affect $W_{ct-1}$ such as fertility rate, etc.

In the two-stage design, I further suspect that there exists some endogeneity in $W_{ct}$[^11]. Both $CO_2$ emissions and women representation in government are highly correlated with industrialization and development. While $X_{ct}$ can eliminate some of the endogeneity of $W_{ct}$ by including economic controls, these covariates alone are not enough to explain a country's development and how it interacts with how women are elected into national legislatures. The rate at which a country develops, both industrially and politically, in regards to the countries in this study may be highly related to which European power originally occupied the territory and thus, how the power left the state when occupation was ceased, including the political institution(s) established. Endogeneity may also arise in what natural resources are available to the country at year $t$. Countries with a greater propensity to use carbon-intensive resources will thus have, on average, higher $CO_2$ emissions. If this also impacts women disproportionately such that women in these countries are more compelled to run for office and enact effective climate policies, then the estimand for $\delta$ will be biased in the negative direction.

[^11]: Mathematically, with endogeneity in the treatment variable, I assume $\exists \quad c_0, t_0 \quad s.t. \quad \mathbb{E}[W_{c_0t_0-1}\epsilon_{c_0t_0}]\neq0$.

To eliminate this endogeneity I invoke a two-stage difference-in-difference design, using an instrumental variable for $W_{ct-1}$. A common instrument in the literature for women parliament has been to use the number of years since the country granted suffrage (Mavisakalyan & Tarverdi, 2019). I use similar instrumental variable design here. Let $Z_{ct-1}$ be the number years since a country $c$ has granted suffrage (measured at year $t$, and lagged one year to align with $W_{ct}$).

In the difference-in-difference design, I use 62 country-fixed effects (omitting the Qatar and Iraq dummies due to colinearity). After lagging $W_{ct}$ one year, I am left with data spanning 1998-2022, thus, including year-fixed effects, I also omit 1998 for colinearity. I utilize the following two-stage estimation approach:
$$
(1) \quad W_{ct-1}=\pi_0+\lambda Z_{ct-1}+\sum_{p=2}^PX_{ct-p}'\Psi_p+
$$
$$
\sum_{k=\text{Angola }k\neq\text{Qatar, Iraq}}^{Zimbabwe}\mu_k\mathbb{1}(\text{Country}_c=k)+\sum_{j=1999}^{2022}\eta_j\mathbb{1}(\text{Year}_t=j)+v_{ct}
$$
$$
(2) \quad Y_{ct}=\beta_0+\delta W_{ct-1}+\sum_{p=2}^PX_{ct-p}'\Omega_p+
$$
$$
\sum_{k=\text{Angola; }k\neq\text{Qatar, Iraq}}^{Zimbabwe}\beta_k\mathbb{1}(\text{Country}_c=k)+\sum_{j=1999}^{2022}\gamma_j\mathbb{1}(\text{Year}_t=j)+\epsilon_{ct}
$$

## The Validity of Years Since Suffrage as an Instrument

Mavisakalyan and Tarverdi, although make use of a cross-sectional data set, to their credit, report that, "an IV estimation can be potentially more convincing in a dataset that varies over time" (161). This is precisely what I employ here. I will expand on the research that uses years since suffrage as an instrument (Grier and Maldonado, 2015; Hicks et al., 2016). Relevance of $Z_{ct-1}$ is convincing with the idea that women in countries where suffrage is granted will see more involvement in government over time as women's rights expand in politics. Mavisakalyan and Tarverdi note that "a country’s history of suffrage should be highly relevant for female exposure to politics" (160).
```{r partial-effects-relevance, echo=F}
women.reg <- read_dta("Data/regression_matrix.dta")

#Partial out fixed effects and covariates
X.covariates = colnames(women.reg)[str_detect(colnames(women.reg), "lgdp|lpopulation")]
fixed.effects = colnames(women.reg)[str_detect(colnames(women.reg), "_I")]

women.reg.cleaned <- women.reg[complete.cases(women.reg[c(X.covariates,
                                       fixed.effects,
                                       "womenrep_lag", "z_lag")]),]

X.covariates = as.matrix(women.reg.cleaned[X.covariates])
fixed.effects = as.matrix(women.reg.cleaned[fixed.effects])

X <- cbind(1, X.covariates, fixed.effects)

Y <- women.reg.cleaned$womenrep_lag

#Partial out Y
Y.tilde <- X%*%(solve(t(X)%*%X)%*%t(X)%*%Y)-Y

#Partial out Z
Z <- women.reg.cleaned$z_lag
Z.tilde <- X%*%(solve(t(X)%*%X)%*%t(X)%*%Z)-Z

#Partialled-out regression on Z
z.lm <- lm(Y.tilde ~ 0+Z.tilde)
z.beta = coef(z.lm)[1]
z.se = summary(z.lm)$coefficients[1,2]
z.ci = (z.beta + c(-1,1)*z.se*qnorm(0.975)) %>% round(5)
```
By partialling-out $Z_{ct-1}$ on the fixed-effects and control variables[^12], the regression coefficient is significantly positive: `r z.beta %>% round(5)` (`r z.ci[1]`, `r z.ci[2]`). Hence, we have significant evidence to believe in the relevance condition of using (lagged) years since suffrage as an instrument.

[^12]: This is the coefficient and 95% C.I. on $\lambda$ from the first stage regression. See Figures 3 and 4 in the Appendix to visually assess the relevance condition of $Z_{ct-1}$.

Regardless of when a country granted suffrage, I assert that the potential carbon emissions per capita as function of the proportion of women in the national legislature in a country $c$ are only dependent upon $W_{ct-1}$ after fixed effects are accounted for, and not on $Z_{ct-1}$. Mathematically, $Z_{ct-1} \perp Y_{ct}(W_{ct-1})|X_{ct-1}, \text{Country}_c, \text{Year}_t$. However, as Mavisakalyan & Tarverdi note, "[our instrument] would be questionable if there is the possibility that the history of suffrage affects climate change policy outcomes through mechanisms other than the Female in Parliament." To satisfy exclusion, I assume,
$$
(3) \quad \mathbb{E}[Z_{ct-1}v_{ct}|\mathbb{X}_{ct}]=0\quad\forall c,t
$$
Where $\mathbb{X}_{ct}$ is the matrix of covariates (including country and yearly fixed effects) I include in equations (1) and (2). This could be violated if countries whose governments are established after colonial rule have (i) strong dependencies to the formal colonial power in terms of institutional structure and (ii) as a function of that institutional structure, relies on carbon-intensive resources to kick start its economy in post-colonial development. In other words, the instrument will be invalid if there are interactions between time, country, and carbon emissions. More research needs to be done to see if this is the case, however, countries that granted suffrage to women immediately after occupation typically took place in the late 1900s, if it was less than random which countries these were, then the economic development---in relation to carbon-intensive resources---is also most likely non-random as well and thus would cause endogeneity in the instrument[^13]. In this paper, I assume away this endogeneity by assuming there is no distinct correlation between when a country granted suffrage and the types of resources the country typically consumes. Critically, I assume monotonicity: the longer a country has passed from the year it granted suffrage, the more, on average, women are elected into national legislatures. I assume that there are no countries for which this trend reverses---though there may be countries that always elect women at high rates (always-takers) or always elect women at low rates (never-takers) regardless of when suffrage was granted; the fact that the instrument may have a weak effect for some countries is tolerable. The monotonicity assumption may be violated if women become less interested in politics over time for certain countries such that their participation decreases and hence the average proportion of women in national legislatures decrease with respect to years since suffrage but would otherwise increase when a country first grants suffrage[^14]. I assume that trends that reverse in a given year is due to country-specific, year-specific fixed effects and or economic-related included covariates, and not because of the instrument[^15].

[^13]: In order to assert this, one would have to see a distinct pattern between the time periods of when women were granted suffrage and some other metric associated with carbon-intensive resource use such as oil exports.
[^14]: Additionally, monotonicity may be violated if obstacles to women's ability to participate in democratic elections become more severe over time.
[^15]: It is also possible that even after controlling for country and year fixed effects along with economic covariates, there could exist a non-parametric (a non-linear and non-monotonic) relationship between our instrument and treatment variable: i.e. when suffrage is granted in a country, voting interest is high, overtime that decreases, and for some reason it reverses course with quadratic relationship and increases again. Equation (1) would not capture such a relationship and the monotonicity assumption would fail.

## Assumptions for Difference-in-Differences

In the absence of more (or less) women being elected conditioned on the proportion of women elected being influenced by the instrument of years since suffrage was granted in the country, the total $CO_2$ emissions per capita for each country would have followed the same trend over time as the other countries that did not elect more (or less) women in national legislatures at that specific time. Since my treatment variable ($W_{ct-1}$) is continuous, the parallel trends assumption is hard to verify or even visually asses. However, I assume no other events or factors are influencing $CO_2$ emissions per capita differently across different countries over time. I assume no unobserved heterogeneity; my estimate is robust to omitted covariates---this is most likely satisfied if we believe in the exclusion restriction of our instrumental variable (equation (3)).

Following Bertrand et. al (2004), I cluster standard errors by countries, where I assume standard errors across countries are independent from each other. This is most appropriate given that within each country, $CO_2$ emissions are most likely correlated with each other given that the economic and industrial structure mostly remains consistent within each country from year to year (yet different across country). This assumption could be violated if groups of countries form specific treaties and agreements to reduce $CO_2$ emissions or make other agreements pertaining to usage of other carbon-intensive resources.

# Results

The difference-in-difference parameter, $\delta$, was estimated using a two-stage design and assumed to be linear[^16]. If my assumptions hold, then the economic and statistic implications are relatively straightforward (interpreted on a 1% scale): On average, for every 1% the average proportion of women in national legislatures in African and Arab countries increase by, the total yearly (gigatons of) $CO_2$ emissions per capita decrease by 1.08 (-1.64, -0.51) (significant at the 5% level). Interestingly, the only significant economic indicator is the first lag of GDP (lagged 2 years).

[^16]: See Table 3 in the Appendix for the regression results and 95% confidence intervals.

# Placebo Regression

To assess the significance of my results, I run a placebo regression by first theorizing a priori a new vector in place of $CO_2$ emissions per capita that in theory, wouldn't be affected by the proportion of women in the national legislature in a given country in a given year[^17]. I used data from the Centre for Research on the Epidemiology of Disasters (CRED) to measure the number natural disasters that occur within a country within a given year. Although there is significant evidence, particularly from the IPCC, to suggest that the rate and severity of natural disasters has been increasing over time, using my identification strategy (equations (1) and (2)), year-fixed effects should account for any trends in time and country-fixed effects should account for countries particularly vulnerable to natural disasters[^18]. Although, intrinsically, the number of natural disasters that occur within a country at year $t$ seems poisson-distributed[^19], however, upon partialling out the regressand for fixed effects, the response variable could be reasonably modeled as a Normal random vector[^20]. The goal of this placebo regression is to reason that there really is signal in the treatment on per capita $CO_2$ emissions. By using my identification strategy outlined previously, I hope to show that mechanically, there is something within the (local) causal chain that influences $CO_2$ emissions to decrease when average proportion of women in national legislatures increase. If the gender composition of national legislatures isn't influencing the number of natural disasters that occur within a country, this supports the theory that I did not obtain the regression results in the original regression by random chance. Indeed, after running my regression, the placebo regression results support this. See Table 4 in the Appendix.

[^17]: This is a non-trivial task from a social science perspective. If women elected within these countries during this time span are actually given power to enact change (and this paper takes the stance that women are given this opportunity and power if elected), then nationally elected officials who are women may potentially affect everything within an economy and government.

[^18]: Reasonably, one could also make the argument that with the increased rate and severity of natural disasters, the propensity for women to enact effective climate policy may increase, and thus, if women within a particular country are able to pass effective climate policies to reduce the negative externalities seen from climate change, one would expect either (i) a negative coefficient for estimation on $\delta$, or conversely (ii) reverse causality, where increased disasters may increase the average proportion of women in national legislatures, if women (and their constituents) are driven by this issue. However, as discussed in the identification portion of this paper, I hope to eliminate most of the endogeneity (including the potential for reverse causality) with my identification strategy. Using a two-stage estimation approach with lags implies that any change in $Y_{ct}$ must come *after* a change in $W_{ct}$ by construction.
[^19]: See Figures 5 and 6 in the Appendix. The placebo regression was also estimated using a iterative generalized methods of moments (GMM) framework by specifying $\mathbb{E}[Z_{ct-1}(Y_{ct-1}-exp\{\mathbb{X}_{ct}'[\beta_0|\delta|\Omega_{1}|...|\Omega_{4}|\beta_{\text{Angola}}|...|\beta_{\text{Zimbabwe}}|\gamma_{1999}|...|\gamma_{2022}]'\})]=0$ as the relevant moment condition for instrumental variables poisson regression. Upon estimation, however, The estimated Jacobian matrix was highly singular so the coefficients and standard errors did not converge.
[^20]: Regardless, the GMM asymptotic variance (used for two-stage least squares) converges in distribution to a Normal distribution even upon clustering (I cluster on country here as well).

## Economic Implications for Climate Policy

My estimation strategy identifies a local average treatment effect on how increases in average proportions of women in national legislatures in African and Arab countries affect yearly per capita $CO_2$ emissions. In this empirical example, the local average treatment effect applies to countries that see greater political engagement[^21] (and thus, a greater proportion of women in national legislature) among women the more progressed they are after suffrage is granted. Given that $\hat{\delta}$ is significantly negative, we have reason to believe that within countries (and years) that elect more women into government as a function of being more politically developed (that is, indicated by years since suffrage was given), per capita $CO_2$ emissions significantly fall at the end of the following year. Among these implications, it is also interesting that the GDP from two years prior (or equivalently, one year prior to when women are elected into office), is significantly positive, indicating that either (i) economic growth facilitates higher probabilities for women being elected into office (supporting my thesis for a causal chain of inference), or (ii) use of carbon-intensive resources may promote economic growth in terms of GDP, and, if externalities placed on women are significant, encourage women to run for office to alleviate the very same externalities that facilitated the economic growth in the first place. Although if equation (3) holds, then the instrumental variable may successfully eliminate entirely the plausibility of reverse causality. Alternatively, contrary to my initial assumption, successful climate policies may not hinder economic growth at all. It could be that that that the policies that are being passed to lower $CO_2$ emissions are also catalysts of promotions for substitute goods such as subsidies for alternative energy sources. These projects could lower $CO_2$ emissions while boosting a country's GDP. While this paper cannot investigate specific interactions between a country's economy, gender composition in national government, and $CO_2$ emissions, it seems clear (with high confidence), when more women enter the national legislature, we see lower per capita $CO_2$ emissions.

[^21]: Notably, since the instrument is continuous, more heterogeneity in the treatment variable is tolerable.

# Conclusion

This paper reaffirms the effect that was outlined by Mavisakalyan and Tarverdi. Utilizing a two-stage difference-in-difference econometric framework, as national legislatures in African and Arab nations witness higher concentrations of women we see significantly lower per capita $CO_2$ emissions. While the exact mechanism through which women's presence influences climate policy remains open for further investigation, these findings hold significant policy implications. By encouraging women to run for office and dismantling barriers to their political participation, governments can potentially achieve environmental benefits alongside broader societal advancements. While this study focused on the local average treatment effect across countries, future research should dive deeper into specific economic policy interactions and its implications on women in government and climate change. Furthermore, investigating the specific policies enacted by legislatures with higher female representation can shed light on the mechanisms at play. This study underscores the importance of considering gender equality within climate change mitigation strategies. Understanding how women leadership shapes environmental policy is crucial for developing more effective strategies to achieve a sustainable future.

\newpage

# Appendix

# Tables & Figures

```{r visuals, echo=F, warning=F, message=F, fig.align='center', fig.width=16, fig.height=16}
label_data <- women %>%
  filter(year >= 1997) %>%
  group_by(countrycode) %>%
  summarize(
    ryear = sample(min(year):max(year), 1),
    womenrep = womenrep[which(year == ryear)],
    co2 = co2[which(year == ryear)]
  )

women %>% filter(year >= 1996) %>%
  ggplot(aes(x=year, y=womenrep, group=country))+
  geom_line()+
  # geom_text_repel(
  #   data = label_data,
  #   aes(x=ryear, label = countrycode),
  #   label.padding = 0.25,
  #   box.padding = unit(0.5, "lines"),  # Padding around the box
  #   box.color = "black",  # Box color
  #   nudge_y = 0.1,  # Adjust position of labels
  #   direction = "both",
  #   max.overlaps = 20,
  #   segment.size = 0.4,
  #   nudge_x = 0.3,
  #   color="black"
  # ) +
  labs(
    x="Year",
    y="Proportion of Women in Government",
    title="Figure 1 - Average Proportion of Women in National Legislatures Over Time"
  )+
  guides(color = "none") +
  theme(axis.text = element_text(size = 36))+
  facet_wrap(.~country, nrow=8)+
  scale_x_continuous(labels = function(x) substr(x, 3, 4))+
  theme_minimal()

women %>% filter(year >= 1996) %>%
  ggplot(aes(x=year, y=co2, group=country))+
  geom_line()+
  # geom_text_repel(
  #   data = label_data,
  #   aes(x=ryear, label = countrycode),
  #   label.padding = 0.25,
  #   box.padding = unit(0.5, "lines"),  # Padding around the box
  #   box.color = "black",  # Box color
  #   nudge_y = 0.3,  # Adjust position of labels
  #   direction = "both",
  #   segment.size = 0.4,
  #   nudge_x = 0.1,
  #   max.overlaps = 20,
  #   color="black"
  # ) +
  guides(color = "none") +
  labs(
    x="Year",
    y=expression("Gigatons of "~CO[2]~" Emissions Per Capita"),
    title="Figure 2 - Per Capita Carbon Emissions Over Time"
  )+
  facet_wrap(.~country, nrow=8, scales = "free_y")+
  scale_x_continuous(labels = function(x) substr(x, 3, 4))+
  theme(axis.text = element_text(size = 36))+
  theme_minimal()
```

```{r summary_stats, echo=F}
women.subset <- women %>% filter(year >= 1997)

tibble(
  Variable = c("CO2 Emissions Per Capita", "Proportion of Women in Gvt", "ln(GDP)", "ln(Population)", "Suffrage"),
  Mean = c(mean(women.subset$co2, na.rm = T) %>% round(3) %>% format(scientific=F), mean(women.subset$womenrep, na.rm=T) %>% round(3) %>% format(scientific=F), mean(log(women.subset$gdp), na.rm=T) %>% round(), mean(log(women.subset$population), na.rm=T) %>% round(3) %>% format(scientific=F),
           mean(women.subset$suffrage, na.rm=T) %>% round(3) %>% format(scientific=F)),
    SD = c(sd(women.subset$co2, na.rm = T) %>% round(3) %>% format(scientific=F), sd(women.subset$womenrep, na.rm=T) %>% round(3) %>% format(scientific=F),
           sd(log(women.subset$gdp), na.rm=T) %>% round(1), sd(log(women.subset$population), na.rm=T) %>% round(3) %>% format(scientific=F),
           sd(women.subset$suffrage, na.rm=T) %>% round(3) %>% format(scientific=F)),
    Min = c(min(women.subset$co2, na.rm = T) %>% round(3) %>% format(scientific=F), min(women.subset$womenrep, na.rm=T) %>% round(3) %>% format(scientific=F), min(log(women.subset$gdp), na.rm=T) %>% round(), min(log(women.subset$population), na.rm=T) %>% round(3) %>% format(scientific=F),
           min(women.subset$suffrage, na.rm=T) %>% round(3) %>% format(scientific=F)),
    Max = c(max(women.subset$co2, na.rm = T) %>% round(3) %>% format(scientific=F), max(women.subset$womenrep, na.rm=T) %>% round(3) %>% format(scientific=F), max(log(women.subset$gdp), na.rm=T) %>% round(1), max(log(women.subset$population), na.rm=T) %>% round(3) %>% format(scientific=F),
           max(women.subset$suffrage, na.rm=T) %>% round(3) %>% format(scientific=F)),
) %>% knitr::kable(caption = "Summary Statistics of Key Variables")

women.subset %>%
  group_by(country) %>%
  summarize(
    N = sum(!is.na(womenrep)),
    `CO2 Emissions Per Capita` = mean(co2, na.rm=T) %>% round(5),
    `Proportion of Women in Gvt` = mean(womenrep, na.rm=T) %>% round(5),
    GDP = mean(gdp, na.rm = T) %>% round(),
    Population = mean(population, na.rm=T) %>% round(3),
    Suffrage = first(suffrage)
  ) %>% knitr::kable(col.names = c("Country", "N", "Mean CO2 Emissions Per Capita", "Mean Proportion of Women in Gvt", "Mean GDP", "Mean Population", "Suffrage"),
                     caption="Summary Statistics by Country") %>% kableExtra::kable_minimal() %>% kableExtra::kable_styling(font_size = 6)
```

```{r echo=F, message=F, warning=F, fig.width=10, fig.height=6, relevance}
women %>%
  ggplot(aes(y=lag(womenrep), x=lag(z)))+
  geom_point(size=0.5)+
  geom_jitter(size=0.5, width=0.1, height=0.01)+
  geom_smooth(method = "lm", se=F, color="#44CCAA")+
  theme_minimal()+
  labs(
    x=expression(Z[ct-1]),
    y=expression(W[ct-1]),
    title = "Figure 3 - Relevance of Instrumental Variable",
    caption = expression("Both the instrument and response variable are graphed using the actual raw vectors; this is useful for intuition, but not necessarily a good estimation for "~lambda)
  )


data.frame(x=Z.tilde, y=Y.tilde) %>%
  ggplot(aes(y=y, x=x))+
    geom_point(size=0.5)+
    geom_jitter(size=0.5, width=0.1, height=0.01)+
    geom_smooth(method = "lm", se=F, color="#44CCAA")+
    theme_minimal()+
    labs(
      x=expression(~tilde(Z)[ct-1]),
      y=expression(~tilde(W)[ct-1]),
      title = "Figure 4 - Relevance of Instrumental Variable Using Partialled-out Regressor and Regressand"
    )
```

```{r regression-results, fig.align='center', echo=F, warning=F, results='asis'}
cat(readLines("regression_results.tex"), sep = "\n")
```

```{r n-disasters, warning=F, echo=F, fig.width=16, fig.height=16}
disasters = read_dta("Data/disasters_regression_matrix.dta")

disasters %>% ggplot() +
  geom_histogram(mapping = aes(x=n_disasters), binwidth = 1, fill="#44CCAA", alpha=0.5)+
  labs(
    x="Number of Natural Disasters per Year",
    y="Frequency",
    title="Figure 5 - Distribution of Placebo Random Vector"
  )+
  scale_x_continuous(breaks = 0:10)+
  theme_minimal()

disasters %>% filter(year >= 1996) %>%
  ggplot(aes(x=year, y=n_disasters, group=country))+
  geom_line()+
  guides(color = "none") +
  labs(
    x="Year",
    y=expression("Number of Natural Disasters"),
    title="Figure 6 - Number of Natural Disasters Over Time"
  )+
  facet_wrap(.~country, nrow=8, scales = "free_y")+
  scale_x_continuous(labels = function(x) substr(x, 3, 4))+
  theme(axis.text = element_text(size = 36))+
  theme_minimal()
```

```{r placebo-results, fig.align='center', echo=F, warning=F, results='asis'}
cat(readLines("placebo_results.tex"), sep = "\n")
```

\newpage

# References

Bertrand, M., Duflo, E., & Mullainathan, S. (2004). How much should we trust differences-in-differences estimates? [https://www.nber.org/system/files/working_papers/w31063/w31063.pdf](https://www.nber.org/system/files/working_papers/w31063/w31063.pdf)

Centre for Research on the Epidemiology of Disasters (CRED). (2023). EM-DAT: The Emergency Events Database [database]. Retrieved [03.29.2024], from [https://www.emdat.be/](https://www.emdat.be/)

Emissions Database for Global Atmospheric Research (EDGAR). (n.d.). Edgar online. Retrieved from [https://edgar.jrc.ec.europa.eu/](https://edgar.jrc.ec.europa.eu/)

European Commission, Joint Research Centre (JRC). (2023). Emissions gap report 2023. Publications Office of the European Union. doi:10.2760/481442 [https://edgar.jrc.ec.europa.eu/report_2023](https://edgar.jrc.ec.europa.eu/report_2023)

Grier, R., & Maldonado, B. (2015). Electoral experience, institutional quality, and economic development in Latin America. Oxford Development Studies, 43(2), 253-280.

Hicks, D. L., Hicks, J. H., & Maldonado, B. (2016). Women as policy makers and donors: Female legislators and foreign aid. European Journal of Political Economy, 44(2), 118-135. doi: 10.1016/j.europoleco.2015.12.004

Inter-Parliamentary Union (IPU). (n.d.). Women in national parliaments. [Data set]. Retrieved from [https://data.ipu.org/womon-ranking](https://data.ipu.org/womon-ranking)

Mavisakalyan, A., & Tarverdi, Y. (2019). Gender and climate change: Do female parliamentarians make a difference? Global Labor Organization (GLO) Discussion Paper No. 221. [https://www.sciencedirect.com/science/article/pii/S0176268017304500](https://www.sciencedirect.com/science/article/pii/S0176268017304500)

Medlicott, E. (2021, March 8). The five devastating reasons climate change affects women more than men. Retrieved from [https://www.euronews.com/green/2021/11/09/the-five-devastating-reasons-climate-change-affects-women-more-than-men](https://www.euronews.com/green/2021/11/09/the-five-devastating-reasons-climate-change-affects-women-more-than-men)

Mirziyoyeva, N., & Salahodjaev, D. (2023). Does representation of women in parliament promote economic growth? Considering evidence from Europe and Central Asia. Frontiers in Political Science, 4(2). [DOI: 10.3389/fpos.2023.1203221] [https://www.frontiersin.org/articles/10.3389/fpos.2023.1120287/full](https://www.frontiersin.org/articles/10.3389/fpos.2023.1120287/full)

Schaeffer, Katherine (2020, October 5). Pew Research Center. Key facts about women's suffrage around the world, a century after U.S. ratified 19th Amendment. Retrieved April 1, 2024, from [https://www.pewresearch.org/short-reads/2020/10/05/key-facts-about-womens-suffrage-around-the-world-a-century-after-u-s-ratified-19th-amendment/](https://www.pewresearch.org/short-reads/2020/10/05/key-facts-about-womens-suffrage-around-the-world-a-century-after-u-s-ratified-19th-amendment/)

Sellers, S. (2016). Gender and climate change: A closer look at existing evidence. Global Gender and Climate Alliance (GGCA). [https://wedo.org/gender-and-climate-change-a-closer-look-at-existing-evidence-ggca/](https://wedo.org/gender-and-climate-change-a-closer-look-at-existing-evidence-ggca/)

Skaaning, Svend-Erik; John Gerring; Henrikas Bartusevicius, 2015, "A Lexicial Index of Electoral Democracy", [https://doi.org/10.7910/DVN/29106](https://doi.org/10.7910/DVN/29106), Harvard Dataverse, V6

United Nations Development Programme (UNDP). (2022, March 8). International Women’s Day 2022 “Gender equality today for a sustainable tomorrow”. Retrieved from [https://www.undp.org/speeches/undp-administrators-statement-international-womens-day-2023-8-march](https://www.undp.org/speeches/undp-administrators-statement-international-womens-day-2023-8-march)

U.S. Department of State. (n.d.). Countries and Areas List. Retrieved from [https://www.state.gov/countries-and-areas-list/](https://www.state.gov/countries-and-areas-list/)

The World Bank. (n.d.). World Bank Open Data. Retrieved from [https://databank.worldbank.org](https://databank.worldbank.org)