---
title: "Closing the Gender Gap in Climate"
subtitle: "Can Electing More Women into Office Narrow the Climate Divide?"
author: "Sam Lee"
format: pdf
header-includes:
  - \usepackage{fontspec}
  - \setmainfont{Times New Roman}
  - \usepackage{setspace}
  - \renewcommand{\normalsize}{\fontsize{12}{14.4}\selectfont}
  - \usepackage[margin=1in]{geometry}
editor: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
library(tidyverse)
library(patchwork)
library(stargazer)
library(haven)
library(ggrepel)

women <- read_dta("Data/womenrep.dta")
```

# Abstract

```{=latex}
\begin{doublespace}
```
I seek to identify the causal effect that electing higher proportions of
women in national legislatures in African and Arab nations has on yearly
per capita $CO_2$ emissions. I utilize a two-stage
difference-in-difference approach on 64 different countries throughout
Africa and the Middle East from 1998-2022. To eliminate the potential
endogeneity in using the proportion of women in national legislatures as
a treatment on per capita $CO_2$ emissions, I use years since women were
granted suffrage in each respective country as an instrumental variable.
One nominal paper in the literature has shown evidence that increased
proportions of women in parliaments are more than likely causally related
to stricter climate policies (Mavisakalyan & Tarverdi, 2019). However,
this paper does not take into account how carbon emissions and women in
government have evolved. To my knowledge, little research has
been done to show an intertemporal causal link between women in
parliaments and its effects on climate. I expand on this research by
showing how carbon emissions have changed over time to elicit a local
average treatment effect between changes in gender compositions at the
national legislature level. When the average proportion of women in
national legislatures in African and Arab countries increases by 1%, I
find a significant effect that total yearly (gigatons of) $CO_2$
emissions per capita decrease by 1.08 (-1.64, -0.51), on average. To
enforce the credibility of significance, I run a placebo regression,
using the number of natural disasters[^1] that occur within a country in
a particular year as an effective placebo.

[^1]: As classified and recorded by the Centre for Research on the
    Epidemiology of Disasters (CRED).
```{=latex}
\end{doublespace}
```
# Introduction

```{=latex}
\begin{doublespace}
```
The current climate and economic research today overwhelmingly support
the thesis that women---especially in developing countries---are
disproportionately affected by the negative externalities of climate
change ([A1](#a1)). If women are disproportionately affected by climate
externalities, one would expect that female policymakers would enact
effective climate policy at higher rates relative to other policymakers,
assuming they have the means and power to do so. Thus, the causal
question of interest reasonably follows as to whether increased proportions
of women decrease $CO_2$ emissions. As we will see, the nature of
causality of this question is rather difficult to elicit due to
intricate endogeneities that persist throughout this problem. However,
the focus of this paper falls on two important motives: (1) The
necessity and obligation of answering the question at hand
considering its implications on climate policy, and (2) The econometric
implications of my identification strategy with its attempt to
reasonably arrive at a causal effect[^2]. The outline of this paper will
proceed as follows: I will first introduce the data I collected for this
study and assess its reliability. I will then discuss the identification
strategy I use to estimate the local average treatment effect. I
support this result with a placebo analysis, and then I will finish by
discussing the results of my econometric analysis and its implications, and
then conclude.

[^2]: As I will show in this paper, the causal effect that I identify is a local average treatment effect with regards to the
    instrumental variable that I invoke.
```{=latex}
\end{doublespace}
```
# Data
```{=latex}
\begin{doublespace}
```
The data used for this analysis was strategically gathered using
publicly available sources[^3]. The Inter-Parliamentary Union (IPU)
maintains a consistent archive of records that monitors the number of
women elected into the national legislature for each country. I obtained emissions data, including total yearly
(gigatons of) $CO_2$ emissions per capita per country, from the
Emissions Database for Global Atmospheric Research (EDGAR) as maintained
by the European Commission[^6]. I obtained country-year-level specific
covariates, such as population and GDP, from the World Bank[^7]. I
obtained the database linking each country's year they passed a right
of universal suffrage from "A Lexical Index of Electoral Democracy" on
the Harvard Dataverse (Skaaning et. al, 2015). I checked
country-specific anomalies using research from the Pew Research
Center (Schaeffer, 2020). Finally, I used the U.S. Department of State's
regional definitions to define which countries I define as either
African or Arab.

[^3]: Data was collected using ethical web-scraping methods or obtained
    via direct download. Further documentation and the scripts used to
    obtain the data used for this analysis can be found on this
    project's Github repository
    <https://github.com/SamLeeBYU/Elections>.

[^6]: A graphical illustration of the yearly $CO_2$ emissions over time
    for each country in my sample can be seen in Figure 2 of the
    Appendix.

[^7]: Statistics for these key variables are summarized in Table 1 of
    the Appendix.

```{=latex}
\end{doublespace}
```
```{=latex}
\begin{doublespace}
```
One potential problem with
data is that I do not account for country-specific government
intricacies that could prevent women or a legislature from
passing effective climate policy---such as not accounting for government
corruption or other unique anomalies (wars between countries, coups,
laws that are discriminatory against women, etc, that prevent a
government from acting efficiently)---at every year[^9]. Merging the suffrage data with all the IPU available data and the $CO_2$ emissions data from EDGAR, I was left with 64 countries spanning 1997-2022. Refer to [A2](#a2) for more specific details on the nature of the data. 

[^9]: To assert consistency in my estimates, econometrically, I assume
    that on average, these factors in the idiosyncratic term of my
    econometric model are 0 for any given country at year $t$. (See
    equation (3)).
```{=latex}
\end{doublespace}
```

# Identification

```{=latex}
\begin{doublespace}
```
To estimate the causal effect that electing more women in national
legislatures in African and Arab countries have on reducing per capita
$CO_2$ emissions, I employ a two-stage difference-in-difference
identification strategy. In regards to curbing the negative
externalities of climate change in the context of this paper, I use per
capita $CO_2$ emissions as a response variable. Let $Y_{ct}$ represent
the total (gigatons of) carbon emissions per capita for a country ($c$)
in a year ($t$). I seek to identify consistently and interpret the local
average treatment effect of $\delta$, the effect of higher proportion of
women in national legislatures in African and Arab countries on national
$CO_2$ emissions per capita. Research from Mirziyoyeva and Salahodjaev
(2023) suggests that women in government promote economic growth. This
could pose potential problems of reverse causality: We need to be
certain that any effect we see on women's representation causes a
reduction (or increase) in carbon emissions. If carbon emissions are
highly correlated with economic growth and industrialization then an OLS
regression (that is, using a typical difference-in-differences approach)
on $Y_{ct}$ to estimate $\delta$ would almost certainly be biased and
inconsistent. To first eliminate the endogeneity for reverse
causality---the supposition that perhaps lower carbon emissions cause a
higher proportion of women to become elected during the same year; or
rather, perhaps higher carbon emissions are correlated with other
variables such as economic growth that are linked to higher proportions
of women becoming elected---I lag the treatment variable of interest,
$W_{ct}$, the average proportion of women in national legislature for
country $c$ during year $t$ one year. This implies that any change in
$Y_{ct}$ will inherently come *after* a change in $W_{ct}$. In the two-stage design, I further suspect that there exists some
endogeneity in $W_{ct}$[^13]. Both $CO_2$ emissions and women's
representation in government are highly correlated with
industrialization and development. While I do
not focus on specific regulatory practices within this paper, an a priori assumption could be that effective
climate policy would consist of enacting particular policies and
regulations that would reduce the $CO_2$ emissions by economic
constraint. Hence, I include a sequence of matrices of autoregressive economic controls, represented by $\sum_{p=2}^PX_{ct-p}$ (where $p$ is the $p$th lag of the economic covariate[^12]). I include logged GDP and logged population in $X_{ct-p}$ (refer to [A3](#a3) to find out about why these controls are included in this way). While $X_{ct}$ can eliminate some of
the endogeneity of $W_{ct}$ by including economic controls, these
covariates alone are not enough to explain a country's development and
how it interacts with how women are elected into national legislatures.
The rate at which a country develops, both industrially and politically,
in regards to the countries in this study may be highly related to which
European power originally occupied the territory and thus, how the power
left the state when the occupation ceased, including the political
institution(s) established. Endogeneity may also arise in what natural
resources are available to the country at year $t$. Countries with a
greater propensity to use carbon-intensive resources will thus have, on
average, higher $CO_2$ emissions. If this also impacts women
disproportionately such that women in these countries are more compelled
to run for office and enact effective climate policies, then the
estimand for $\delta$ will be biased in the negative direction.

[^12]: A priori I set $P=5$, the total number of lags on all the economic covariate controls.
[^13]: Mathematically, with endogeneity in the treatment variable, I
    assume
    $\exists \quad c_0, t_0 \quad s.t. \quad \mathbb{E}[W_{c_0t_0-1}\epsilon_{c_0t_0}]\neq0$.
```{=latex}
\end{doublespace}
```

```{=latex}
\begin{doublespace}
```
To eliminate this endogeneity I invoke a two-stage
difference-in-difference design, using an instrumental variable for
$W_{ct-1}$. A common instrument in the literature for women's representation in parliament
has been to use the number of years since the country granted suffrage
(Mavisakalyan & Tarverdi, 2019). I use a similar instrumental variable
design here. Let $Z_{ct-1}$ be the number of years since a country $c$ has
granted suffrage (measured at year $t$, and lagged one year to align
with $W_{ct-1}$). In the difference-in-difference design, I use 62 country-fixed effects
(omitting the Qatar and Iraq dummies due to colinearity). After lagging
$W_{ct}$ one year, I am left with data spanning 1998-2022, thus,
including year-fixed effects, I also omit 1998 for colinearity. I
utilize the following two-stage estimation approach:
```{=latex}
\end{doublespace}
```
$$
(1) \quad W_{ct-1}=\pi_0+\lambda Z_{ct-1}+\sum_{p=2}^PX_{ct-p}'\Psi_p+
$$
$$
\sum_{k=\text{Angola }k\neq\text{Qatar, Iraq}}^{Zimbabwe}\mu_k\mathbb{1}(\text{Country}_c=k)+\sum_{j=1999}^{2022}\eta_j\mathbb{1}(\text{Year}_t=j)+v_{ct}
$$
$$
(2) \quad Y_{ct}=\beta_0+\delta W_{ct-1}+\sum_{p=2}^PX_{ct-p}'\Omega_p+
$$
$$
\sum_{k=\text{Angola; }k\neq\text{Qatar, Iraq}}^{Zimbabwe}\beta_k\mathbb{1}(\text{Country}_c=k)+\sum_{j=1999}^{2022}\gamma_j\mathbb{1}(\text{Year}_t=j)+\epsilon_{ct}
$$

```{r partial-effects-relevance, echo=F}
women.reg <- read_dta("Data/regression_matrix.dta")

#Partial out fixed effects and covariates
X.covariates = colnames(women.reg)[str_detect(colnames(women.reg), "lgdp|lpopulation")]
fixed.effects = colnames(women.reg)[str_detect(colnames(women.reg), "_I")]

women.reg.cleaned <- women.reg[complete.cases(women.reg[c(X.covariates,
                                       fixed.effects,
                                       "womenrep_lag", "z_lag")]),]

X.covariates = as.matrix(women.reg.cleaned[X.covariates])
fixed.effects = as.matrix(women.reg.cleaned[fixed.effects])

X <- cbind(1, X.covariates, fixed.effects)

Y <- women.reg.cleaned$womenrep_lag

#Partial out Y
Y.tilde <- X%*%(solve(t(X)%*%X)%*%t(X)%*%Y)-Y

#Partial out Z
Z <- women.reg.cleaned$z_lag
Z.tilde <- X%*%(solve(t(X)%*%X)%*%t(X)%*%Z)-Z

#Partialled-out regression on Z
z.lm <- lm(Y.tilde ~ 0+Z.tilde)
z.beta = coef(z.lm)[1]
z.se = summary(z.lm)$coefficients[1,2]
z.ci = (z.beta + c(-1,1)*z.se*qnorm(0.975)) %>% round(5)
```
## The Validity of Years Since Suffrage as an Instrument
```{=latex}
\begin{doublespace}
```
Mavisakalyan and Tarverdi, although make use of a cross-sectional data
set, report that, "an IV estimation can be potentially
more convincing in a dataset that varies over time" (161). This is
precisely what I employ here. I will expand on the research that uses
years since suffrage as an instrument (Grier and Maldonado, 2015; Hicks
et al., 2016). Relevance of $Z_{ct-1}$ is convincing with the idea that
women in countries where suffrage is granted will see more involvement
in government over time as women's rights expand in politics.
Mavisakalyan and Tarverdi note that "a country's history of suffrage
should be highly relevant for female exposure to politics" (160). By partialling-out $Z_{ct-1}$ on the fixed-effects and control
variables[^14], the regression coefficient is significantly positive:
`r z.beta %>% round(5)` (`r z.ci[1]`, `r z.ci[2]`). Regardless of when a country granted suffrage, I assert that the
potential carbon emissions per capita as a function of the proportion of
women in the national legislature in a country $c$ are only dependent
upon $W_{ct-1}$ after fixed effects are accounted for, and not on
$Z_{ct-1}$. Mathematically,
$Z_{ct-1} \perp Y_{ct}(W_{ct-1})|X_{ct-1},...,X_{ct-P}, \text{Country}_c, \text{Year}_t$.
However, as Mavisakalyan & Tarverdi note, "\[our instrument\] would be
questionable if there is the possibility that the history of suffrage
affects climate change policy outcomes through mechanisms other than the
Female in Parliament." To satisfy exclusion, I assume, $$
(3) \quad \mathbb{E}[Z_{ct-1}v_{ct}|\mathbb{X}_{ct}]=0\quad\forall c,t
$$ Where $\mathbb{X}_{ct}$ is the matrix of covariates (including
country and yearly fixed effects) I include in equations (1) and (2).
This could be violated if countries whose governments are established
after colonial rule have (i) strong dependencies on the formal colonial
power in terms of institutional structure and (ii) as a function of that
institutional structure relies on carbon-intensive resources to kick
start its economy in post-colonial development. In other words, the
instrument will be invalid if there are interactions between time,
country, and carbon emissions. More research needs to be done to see if
this is the case, however, in countries that granted suffrage to women
immediately after occupation typically took place in the late 1900s, if
it was less than random which countries these were, then the economic
development---in relation to carbon-intensive resources---is also most
likely non-random as well and thus would cause endogeneity in the
instrument[^15]. In this paper, I assume away this endogeneity by
assuming there is no distinct correlation between when a country granted
suffrage and the types of resources the country typically consumes.
Critically, I assume monotonicity: the longer a country has passed from
the year it granted suffrage, the more, on average, women are elected
into national legislatures. I assume that there are no countries for
which this trend reverses[^11]. The monotonicity assumption may be violated if women become
less interested in politics over time for certain countries such that
their participation decreases and hence the average proportion of women
in national legislatures decrease with respect to years since suffrage
but would otherwise increase when a country first grants suffrage[^16].
I assume that trends that reverse in a given year are due to
country-specific, year-specific fixed effects and or economic-related
included covariates, and not because of the instrument[^17].

[^14]: This is the coefficient and 95% C.I. on $\lambda$ from the first
    stage regression. See Figures 3 and 4 in the Appendix to visually
    assess the relevance condition of $Z_{ct-1}$.

[^15]: To assert this, one would have to see a distinct pattern
    between the periods of when women were granted suffrage and other metrics associated with carbon-intensive resource use such
    as oil exports.
    
[^11]: Though there may be countries that always elect women at high rates (always-takers) or always elect women at low rates (never-takers) regardless of when suffrage was granted; the fact that the instrument may have a weak effect for some countries is tolerable.

[^16]: Additionally, monotonicity may be violated if obstacles to
    women's ability to participate in democratic elections become more
    severe over time.

[^17]: It is also possible that even after controlling for country and
    year fixed effects along with economic covariates, there could exist
    a non-parametric (a non-linear and non-monotonic) relationship
    between our instrument and treatment variable: i.e. when suffrage is
    granted in a country, where voting interest is initially high, but overtime that
    decreases, but for some reason, it reverses course with some quadratic or non-parametric
    relationship and increases again. Equation (1) would not capture
    such a relationship and the monotonicity assumption would fail.
```{=latex}
\end{doublespace}
```

## Assumptions for Difference-in-Differences

```{=latex}
\begin{doublespace}
```
In the absence of more (or less) women being elected conditioned on the
proportion of women elected being influenced by the instrument of years
since suffrage was granted in the country, I assume the total $CO_2$ emissions
per capita for each country would have followed the same trend over time
as the other countries that did not elect more (or less) women in
national legislatures at that specific time. Since my treatment variable
($W_{ct-1}$) is continuous, the parallel trends assumption is hard to
verify or even visually asses. I assume no unobserved heterogeneity; my
estimate is robust to omitted covariates---this assumption is satisfied
if we believe in the exclusion restriction of our instrumental variable
(equation (3)). Following Bertrand et. al (2004), I cluster standard errors by
countries, where I assume standard errors across countries are
independent from each other. This is most appropriate given that within
each country, $CO_2$ emissions are most likely correlated with each
other given that the economic and industrial structure mostly remains
consistent within each country from year to year (yet different across
country). This assumption could be violated if groups of countries form
specific treaties and agreements to reduce $CO_2$ emissions or make
other agreements pertaining regarding the usage of carbon-intensive
resources.
```{=latex}
\end{doublespace}
```

# Results

```{=latex}
\begin{doublespace}
```
The difference-in-difference parameter, $\delta$, was estimated using a
two-stage design and assumed to be linear[^18]. If my assumptions hold,
then the economic and statistical implications are relatively
straightforward (interpreted on a 1% scale): On average, for every 1%
the average proportion of women in national legislatures in African and
Arab countries increase by, the total yearly (gigatons of) $CO_2$
emissions per capita decrease by 1.08 (-1.64, -0.51) (significant at the
5% level). Interestingly, the only significant economic indicator is the
first lag of GDP (lagged 2 years).

[^18]: See Table 3 in the Appendix for the regression results and 95%
    confidence intervals.
```{=latex}
\end{doublespace}
```

## Placebo Regression

```{=latex}
\begin{doublespace}
```
To assess the significance of my results, I run a placebo regression by
first theorizing a priori a new vector in place of $CO_2$ emissions per
capita that in theory, wouldn't be affected by the proportion of women
in the national legislature in a given country in a given year[^19]. I
used data from the Centre for Research on the Epidemiology of Disasters
(CRED) to measure the number of natural disasters that occur within a
country within a given year. Although there is significant evidence,
particularly from the IPCC, to suggest that the rate and severity of
natural disasters have been increasing over time, using my identification
strategy (equations (1) and (2)), year-fixed effects should account for
any trends in time and country-fixed effects should account for
countries particularly vulnerable to natural disasters[^20]. Although,
intrinsically, the number of natural disasters that occur within a
country at year $t$ seems Poisson-distributed[^21], however, upon
partialling out the regressand for fixed effects, the response variable
could be reasonably modeled as a Normal random vector[^22]. The goal of
this placebo regression is to reason that there is a signal in the
treatment on per capita $CO_2$ emissions. If the gender composition of national
legislatures isn't influencing the number of natural disasters that
occur within a country, this supports the theory that I did not obtain
the regression results in the original regression by random chance.
Indeed, after running my regression, the placebo regression results
support this. See Table 4 in the Appendix.

[^19]: This is a non-trivial task from a social science perspective. If
    women elected within these countries during this period are given the power to enact change (and this paper takes the
    stance that women are given this opportunity and power if elected),
    then nationally elected officials who are women may potentially
    affect *everything* within an economy and government.

[^20]: Reasonably, one could also make the argument that with the
    increased rate and severity of natural disasters, the propensity for
    women to enact effective climate policy may increase, and thus, if
    women within a particular country are able to pass effective climate
    policies to reduce the negative externalities seen from climate
    change, one would expect either (i) a negative coefficient for
    estimation on $\delta$, or conversely (ii) reverse causality, where
    increased disasters may increase the average proportion of women in
    national legislatures, if women (and their constituents) are driven
    by this issue. However, as discussed in the identification portion
    of this paper, I hope to eliminate most of the endogeneity
    (including the potential for reverse causality) with my
    identification strategy. Using a two-stage estimation approach with
    lags implies that any change in $Y_{ct}$ must come *after* a change
    in $W_{ct}$ by construction.

[^21]: See Figures 5 and 6 in the Appendix. The placebo regression was
    also estimated using an iterative generalized method of moments
    (GMM) framework by specifying
    $\mathbb{E}[Z_{ct-1}(Y_{ct-1}-exp\{\mathbb{X}_{ct}'[\beta_0|\delta|\Omega_{1}|...|\Omega_{4}|\beta_{\text{Angola}}|...|\beta_{\text{Zimbabwe}}|\gamma_{1999}|...|\gamma_{2022}]'\})]=0$
    as the relevant moment condition for instrumental variables Poisson
    regression. Upon estimation, however, The estimated Jacobian matrix
    was highly singular so the coefficients and standard errors did not
    converge.

[^22]: Regardless, the GMM asymptotic variance (used for two-stage least
    squares) converges in distribution to a Normal distribution even
    upon clustering (I cluster by country in this regression as well).
```{=latex}
\end{doublespace}
```

## Economic Implications for Climate Policy

```{=latex}
\begin{doublespace}
```
My estimation strategy identifies a local average treatment effect on
how increases in average proportions of women in national legislatures
in African and Arab countries affect yearly per capita $CO_2$ emissions.
In this empirical example, the local average treatment effect applies to
countries that see greater political engagement[^23] (and thus, a
greater proportion of women in national legislatures) among women the
more progressed they are after suffrage is granted. Given that
$\hat{\delta}$ is significantly negative, we have reason to believe that
within countries (and years) that elect more women into government as a
function of being more politically developed (that is, indicated by
years since suffrage was given), per capita $CO_2$ emissions
significantly fall at the end of the following year. Among these
implications, it is also interesting that the GDP from two years prior
(or equivalently, one year prior to when women are elected into office),
is significantly positive, indicating that either (i) economic growth
facilitates higher probabilities of women being elected into office
(supporting my thesis for a causal chain of inference ([A3](#a3))), or (ii) use of
carbon-intensive resources may promote economic growth in terms of GDP,
and, if externalities placed on women are significant, encourage women
to run for office to alleviate the very same externalities that
facilitated the economic growth in the first place. Although if equation
(3) holds, then the instrumental variable may successfully eliminate
entirely the plausibility of reverse causality. Alternatively, contrary
to my initial assumption, successful climate policies may not hinder
economic growth at all. It could be that that that the policies that are
being passed to lower $CO_2$ emissions are also catalysts of promotions
for substitute goods such as subsidies for alternative energy sources.
These projects could lower $CO_2$ emissions while boosting a country's
GDP. While this paper cannot investigate specific interactions between a
country's economy, gender composition in the national government, and $CO_2$
emissions, it seems clear (with high confidence) that when more women enter
the national legislature, we see lower per capita $CO_2$ emissions.

[^23]: Notably, since the instrument is continuous, more heterogeneity
    in the treatment variable is tolerable.
```{=latex}
\end{doublespace}
```

# Conclusion

```{=latex}
\begin{doublespace}
```
Utilizing a two-stage difference-in-difference econometric
framework, as national legislatures in African and Arab nations witness
higher concentrations of women we see significantly lower per capita
$CO_2$ emissions. While the exact mechanism through which women's
presence influences climate policy remains open for further
investigation, these findings hold significant policy implications. By
encouraging women to run for office and dismantling barriers to their
political participation, governments can potentially achieve
environmental benefits alongside broader societal advancements. While
this study focused on the local average treatment effect across
countries, future research should dive deeper into specific economic
policy interactions and their implications on women in government and
climate change. This study underscores the importance of considering
gender equality within climate change mitigation strategies.
Understanding how women's leadership shapes environmental policy is
crucial for developing more effective strategies to achieve a
sustainable future.
```{=latex}
\end{doublespace}
```

\newpage
# Appendix

## A1 - Existing Research on Disproportionate Impacts of Climate Change on Women {#a1}
```{=latex}
\begin{doublespace}
```
During periods of increased droughts, which are exacerbated due to the
effects of climate change, women often make long trips to meet
agricultural, hygiene, and family needs. This has been linked to
increased violence against women (UNDP, 2019). Increased risks of
poverty and food insecurities have also been shown to increase as a result
of increased rates of natural disasters and variable weather patterns.
One article reports that "when a family is faced with the impact of the
climate crisis, girls' education is one of the first things families
drop" (Medlicott, 2021). Another meta-analysis consisting of 53 studies
report that two-thirds of those studies find that women are more often
the victims of death or injury in the case of extreme weather events
(Seller, 2016).
```{=latex}
\end{doublespace}
```
## A2 - Details on Data Collection {#a2}
```{=latex}
\begin{doublespace}
```
The IPU Data: Not every government maintains the same structure of government, however, the IPU
distinguishes between women elected into the lower house and upper house
levels of government throughout multiple periods. I take the
average proportion of women holding office in the national legislature
overall in a given year by taking a weighted average of the two
houses. Econometrically, this assumes that on average, the power that women are given in each house of national legislature is equal to the 'weight' of women in that chamber of legislature: In other
words, assuming a fair political system, I assume no chamber can systemically dominate the other chamber. A graphical illustration of the average proportion of women in the national legislatures for each country in my sample can be seen in Figure 1.

Suffrage Data: Not all countries in Africa and the Middle East treat women equally. Even when suffrage is granted, it is not granted universally. Thus, the instrument may not influence the proportion of women in government the same in every country. This does not discredit my econometric analysis, however, so long as the monotonicity assumption of my instrument is met. However, I exclude obvious anomalies in my analytic sample. See Table 2 for summary statistics for each country in my sample. In my sample, the United Arab Emirates (UAE) and Saudia Arabia are excluded because they don't give full suffrage to women: The UAE held
its first national elections in 2006, but only a select amount of men and women were eligible to participate (Schaeffer, 2020). Saudi Arabia doesn't hold national elections. Women are, however, enfranchised at the local level government levels. This leaves us with a total of 64 different countries in the analytic sample.

Notably, the data is left as an unbalanced panel data set (Refer to Table 2). The World Bank does not have GDP or population data for every single year of every single country in the data, and critically, the IPU also notes that it cannot reliably retrieve the number of women in national legislatures for every year. These years for those specific countries are naturally excluded. I assume that the probability at which data for a specific year-country combination will be missing is essentially random. This assumption may be violated if countries that are systemically difficult to obtain data for (such as Somalia or the Central African Republic) is also correlated with other covariates and or $CO_2$ emissions.
```{=latex}
\end{doublespace}
```

## A3 - Causal Chain of Inference {#a3}
```{=latex}
\begin{doublespace}
```
It could be that increased proportions of women in government could
(causally) affect economic performance within a country at time $t$.
Thus, to establish the causal chain of inference, I infer:
$$
\sum_{p=2}^PX_{ct-p}\rightarrow W_{ct-1}\rightarrow Y_{ct}
$$
Where $X_{ct-p}$ is a matrix of controls related to economic growth.
For the same reason the $W_{ct}$ is lagged one year to ensure that a
change in $Y_{ct}$ comes after a change in $W_{ct}$, I include a series
of autoregressive lags in the economic controls (starting at lag 2) to
ensure that changes in $W_{ct-1}$ occur *after* changes in economic
development. A priori, I set $P=5$ in my econometric model to control
for economic lags in a country. For this two-stage
regression, $X_{ct-p}$ consists of logged GDP and logged population for
a country at year $t$. A further analysis could (and should) include further economic
controls that may affect $W_{ct-1}$ such as fertility rate, etc.
```{=latex}
\end{doublespace}
```
# Tables & Figures
```{r summary_stats, echo=F}
women.subset <- women %>% filter(year >= 1997)

tibble(
  Variable = c("CO2 Emissions Per Capita", "Proportion of Women in Gvt", "ln(GDP)", "ln(Population)", "Suffrage"),
  Mean = c(mean(women.subset$co2, na.rm = T) %>% round(3) %>% format(scientific=F), mean(women.subset$womenrep, na.rm=T) %>% round(3) %>% format(scientific=F), mean(log(women.subset$gdp), na.rm=T) %>% round(), mean(log(women.subset$population), na.rm=T) %>% round(3) %>% format(scientific=F),
           mean(women.subset$suffrage, na.rm=T) %>% round(3) %>% format(scientific=F)),
    SD = c(sd(women.subset$co2, na.rm = T) %>% round(3) %>% format(scientific=F), sd(women.subset$womenrep, na.rm=T) %>% round(3) %>% format(scientific=F),
           sd(log(women.subset$gdp), na.rm=T) %>% round(1), sd(log(women.subset$population), na.rm=T) %>% round(3) %>% format(scientific=F),
           sd(women.subset$suffrage, na.rm=T) %>% round(3) %>% format(scientific=F)),
    Min = c(min(women.subset$co2, na.rm = T) %>% round(3) %>% format(scientific=F), min(women.subset$womenrep, na.rm=T) %>% round(3) %>% format(scientific=F), min(log(women.subset$gdp), na.rm=T) %>% round(), min(log(women.subset$population), na.rm=T) %>% round(3) %>% format(scientific=F),
           min(women.subset$suffrage, na.rm=T) %>% round(3) %>% format(scientific=F)),
    Max = c(max(women.subset$co2, na.rm = T) %>% round(3) %>% format(scientific=F), max(women.subset$womenrep, na.rm=T) %>% round(3) %>% format(scientific=F), max(log(women.subset$gdp), na.rm=T) %>% round(1), max(log(women.subset$population), na.rm=T) %>% round(3) %>% format(scientific=F),
           max(women.subset$suffrage, na.rm=T) %>% round(3) %>% format(scientific=F)),
) %>% knitr::kable(caption = "Summary Statistics of Key Variables")
```
\newpage
```{r visuals, echo=F, warning=F, message=F, fig.align='center', fig.width=16, fig.height=16}
women.rsamp <- women %>% filter(year >= 1997) %>%
  filter(country %in% sample(unique(country))[1:10])

label_data <- women.rsamp %>%
  group_by(country) %>%
  summarize(
    ryear = sample(min(year):max(year), 1),
    womenrep = womenrep[which(year == ryear)],
    co2 = co2[which(year == ryear)]
  )

aggregate_womenrep <- women.rsamp %>%
  ggplot(aes(x=year, y=womenrep, group=country, color=country))+
  geom_line()+
  geom_text_repel(
    data = label_data,
    aes(x=ryear, label = country),
    label.padding = 0.25,
    box.padding = unit(0.5, "lines"),  # Padding around the box
    box.color = "black",  # Box color
    nudge_y = 0.1,  # Adjust position of labels
    direction = "both",
    max.overlaps = 20,
    segment.size = 0.4,
    nudge_x = 0.3,
    color="black"
  ) +
  labs(
    x="Year",
    y="Proportion of Women in Government",
    title="Figure 1 - Average Proportion of Women in National Legislatures Over Time"
  )+
  guides(color = "none") +
  theme(axis.text = element_text(size = 36))+
  scale_x_continuous(labels = function(x) substr(x, 3, 4))+
  theme_minimal()

aggregate_co2 <- women.rsamp %>%
  ggplot(aes(x=year, y=co2, group=country, color=country))+
  geom_line()+
  geom_text_repel(
    data = label_data,
    aes(x=ryear, label = country),
    label.padding = 0.25,
    box.padding = unit(0.5, "lines"),  # Padding around the box
    box.color = "black",  # Box color
    nudge_y = 0.3,  # Adjust position of labels
    direction = "both",
    segment.size = 0.4,
    nudge_x = 0.1,
    max.overlaps = 20,
    color="black"
  ) +
  guides(color = "none") +
  labs(
    x="Year",
    y=expression("Gigatons of "~CO[2]~" Emissions Per Capita"),
    title="Figure 2 - Per Capita Carbon Emissions Over Time"
  )+
  scale_x_continuous(labels = function(x) substr(x, 3, 4))+
  theme(axis.text = element_text(size = 36))+
  theme_minimal()

#ggsave("Graphs/percapitaco2.png", aggregate_co2, width=14, height=8)
#ggsave("Graphs/womenrep.png", aggregate_womenrep, width=14, height=8)

women %>% filter(year >= 1996) %>%
  ggplot(aes(x=year, y=womenrep, group=country))+
  geom_line()+
  labs(
    x="Year",
    y="Proportion of Women in Government",
    title="Figure 1 - Average Proportion of Women in National Legislatures Over Time"
  )+
  guides(color = "none") +
  theme(axis.text = element_text(size = 36))+
  facet_wrap(.~country, nrow=8)+
  scale_x_continuous(labels = function(x) substr(x, 3, 4))+
  theme_minimal()

women %>% filter(year >= 1996) %>%
  ggplot(aes(x=year, y=co2, group=country))+
  geom_line()+
  guides(color = "none") +
  labs(
    x="Year",
    y=expression("Gigatons of "~CO[2]~" Emissions Per Capita"),
    title="Figure 2 - Per Capita Carbon Emissions Over Time"
  )+
  facet_wrap(.~country, nrow=8, scales = "free_y")+
  scale_x_continuous(labels = function(x) substr(x, 3, 4))+
  theme(axis.text = element_text(size = 36))+
  theme_minimal()
```
```{r country_stats, echo=F}
women.subset %>%
  group_by(country) %>%
  summarize(
    N = sum(!is.na(womenrep)),
    `CO2 Emissions Per Capita` = mean(co2, na.rm=T) %>% round(5),
    `Proportion of Women in Gvt` = mean(womenrep, na.rm=T) %>% round(5),
    GDP = mean(gdp, na.rm = T) %>% round(),
    Population = mean(population, na.rm=T) %>% round(3),
    Suffrage = first(suffrage)
  ) %>% knitr::kable(col.names = c("Country", "N", "Mean CO2 Emissions Per Capita", "Mean Proportion of Women in Gvt", "Mean GDP", "Mean Population", "Suffrage"),
                     caption="Summary Statistics by Country") %>% kableExtra::kable_minimal() %>% kableExtra::kable_styling(font_size = 6)
```
```{r echo=F, message=F, warning=F, fig.width=10, fig.height=6, relevance}
women %>%
  ggplot(aes(y=lag(womenrep), x=lag(z)))+
  geom_point(size=0.5)+
  geom_jitter(size=0.5, width=0.1, height=0.01)+
  geom_smooth(method = "lm", se=F, color="#44CCAA")+
  theme_minimal()+
  labs(
    x=expression(Z[ct-1]),
    y=expression(W[ct-1]),
    title = "Figure 3 - Relevance of Instrumental Variable",
    caption = expression("Both the instrument and response variable are graphed using the actual raw vectors; this is useful for intuition, but not necessarily a good estimation for "~lambda)
  )


relevance.plot <- data.frame(x=Z.tilde, y=Y.tilde) %>%
  ggplot(aes(y=y, x=x))+
    geom_point(size=0.5)+
    geom_jitter(size=0.5, width=0.1, height=0.01)+
    geom_smooth(method = "lm", se=F, color="#44CCAA")+
    theme_minimal()+
    labs(
      x=expression(~tilde(Z)[ct-1]),
      y=expression(~tilde(W)[ct-1]),
      title = "Figure 4 - Relevance of Instrumental Variable Using Partialled-out Regressor and Regressand"
    )

#ggsave("Graphs/relevance.png", relevance.plot, width=14, height=8)

relevance.plot
```
```{r regression-results, fig.align='center', echo=F, warning=F, results='asis'}
cat(readLines("regression_results.tex"), sep = "\n")
```
```{r n-disasters, warning=F, echo=F, fig.width=16, fig.height=16}
disasters = read_dta("Data/disasters_regression_matrix.dta")

placebo.plt <- disasters %>% ggplot() +
  geom_histogram(mapping = aes(x=n_disasters), binwidth = 1, fill="#44CCAA", alpha=0.5)+
  labs(
    x="Number of Natural Disasters per Year",
    y="Frequency",
    title="Figure 5 - Distribution of Placebo Random Vector"
  )+
  scale_x_continuous(breaks = 0:10)+
  theme_minimal()

ggsave("Graphs/placebo.png", placebo.plt, width=14, height=8)

placebo.plt

disasters %>% filter(year >= 1996) %>%
  ggplot(aes(x=year, y=n_disasters, group=country))+
  geom_line()+
  guides(color = "none") +
  labs(
    x="Year",
    y=expression("Number of Natural Disasters"),
    title="Figure 6 - Number of Natural Disasters Over Time"
  )+
  facet_wrap(.~country, nrow=8, scales = "free_y")+
  scale_x_continuous(labels = function(x) substr(x, 3, 4))+
  theme(axis.text = element_text(size = 36))+
  theme_minimal()
```
```{r placebo-results, fig.align='center', echo=F, warning=F, results='asis'}
cat(readLines("placebo_results.tex"), sep = "\n")
```
\newpage
# References

Bertrand, M., Duflo, E., & Mullainathan, S. (2004). How much should we
trust differences-in-differences estimates?
<https://www.jstor.org/stable/25098683>

Centre for Research on the Epidemiology of Disasters (CRED). (2023).
EM-DAT: The Emergency Events Database \[database\]. Retrieved
\[03.29.2024\], from <https://www.emdat.be/>

Emissions Database for Global Atmospheric Research (EDGAR). (n.d.).
Edgar online. Retrieved from <https://edgar.jrc.ec.europa.eu/>

European Commission, Joint Research Centre (JRC). (2023). Emissions gap
report 2023. Publications Office of the European Union.
doi:10.2760/481442 <https://edgar.jrc.ec.europa.eu/report_2023>

Grier, R., & Maldonado, B. (2015). Electoral experience, institutional
quality, and economic development in Latin America. Oxford Development
Studies, 43(2), 253-280.

Hicks, D. L., Hicks, J. H., & Maldonado, B. (2016). Women as policy
makers and donors: Female legislators and foreign aid. European Journal
of Political Economy, 44(2), 118-135. doi:
10.1016/j.europoleco.2015.12.004

Inter-Parliamentary Union (IPU). (n.d.). Women in national parliaments.
\[Data set\]. Retrieved from <https://data.ipu.org/womon-ranking>

Mavisakalyan, A., & Tarverdi, Y. (2019). Gender and climate change: Do
female parliamentarians make a difference? Global Labor Organization
(GLO) Discussion Paper No. 221.
<https://www.sciencedirect.com/science/article/pii/S0176268017304500>

Medlicott, E. (2021, March 8). The five devastating reasons climate
change affects women more than men. Retrieved from
<https://www.euronews.com/green/2021/11/09/the-five-devastating-reasons-climate-change-affects-women-more-than-men>

Mirziyoyeva, N., & Salahodjaev, D. (2023). Does representation of women
in parliament promote economic growth? Considering evidence from Europe
and Central Asia. Frontiers in Political Science, 4(2). \[DOI:
10.3389/fpos.2023.1203221\]
<https://www.frontiersin.org/articles/10.3389/fpos.2023.1120287/full>

Schaeffer, Katherine (2020, October 5). Pew Research Center. Key facts
about women's suffrage around the world, a century after U.S. ratified
19th Amendment. Retrieved April 1, 2024, from
<https://www.pewresearch.org/short-reads/2020/10/05/key-facts-about-womens-suffrage-around-the-world-a-century-after-u-s-ratified-19th-amendment/>

Sellers, S. (2016). Gender and climate change: A closer look at existing
evidence. Global Gender and Climate Alliance (GGCA).
<https://wedo.org/gender-and-climate-change-a-closer-look-at-existing-evidence-ggca/>

Skaaning, Svend-Erik; John Gerring; Henrikas Bartusevicius, 2015, "A
Lexicial Index of Electoral Democracy",
<https://doi.org/10.7910/DVN/29106>, Harvard Dataverse, V6

United Nations Development Programme (UNDP). (2022, March 8).
International Women's Day 2022 "Gender equality today for a sustainable
tomorrow". Retrieved from
<https://www.undp.org/speeches/undp-administrators-statement-international-womens-day-2023-8-march>

U.S. Department of State. (n.d.). Countries and Areas List. Retrieved
from <https://www.state.gov/countries-and-areas-list/>

The World Bank. (n.d.). World Bank Open Data. Retrieved from
<https://databank.worldbank.org>
