---
title: "Closing the Gender Gap in Climate"
subtitle: "Can Electing More Women into Office Narrow the Climate Divide?"
author: "Sam Lee"
format: pdf
header-includes:
  - \usepackage{fontspec}
  - \setmainfont{Times New Roman}
  - \usepackage{setspace}
  - \renewcommand{\normalsize}{\fontsize{12}{14.4}\selectfont}
  - \linespread{2}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)

library(tidyverse)
library(patchwork)
library(stargazer)
library(haven)
library(ggrepel)

women <- read_dta("Data/womenrep.dta")
```

# Abstract

In this paper, I seek to identify what the causal effect of electing higher proportions of women in national legislatures in African and Arab nations is on yearly $CO_2$ per capita emissions. I utilize a two-stage difference-in-difference approach on 64 different countries throughout Africa and the Middle-East. To eliminate the potential endogeneity in using the proportion of women in national legislature as a treatment on per capita $CO_2$ emissions, I use years since women were granted suffrage in each respective country as an instrumental variable. One nominal economic paper has shown evidence that increased women in parliaments is more than likely causally related to stricter climate policies (Mavisakalyan & Tarverdi, 2019). However, this paper fails to take into account how carbon emissions and women in government have evolved over time. In this paper, I expand on this research by showing directly how carbon emissions have changed over time to elicit a causal effect between changes in gender compositions at the national legislature level. When the average proportion of women in national legislatures in African and Arab countries increase by 1%, I find a significant effect that total yearly (gigatons of) $CO_2$ emissions per capita decrease by 0.273 (-0.306, -0.240) on average.

# Introduction

The current climate and economic research today overwhelmingly supports the thesis that women---especially in developing countries---are disproportionately affected by the negative externalities of climate change. During periods of increased droughts, which are exacerbated due to the effects of climate change, women often make long trips to meet agricultural, hygiene, and family needs. This has been linked to increased violence against women (UNDP, 2019). Increased risks of poverty, food insecurities have also been shown to increase as a result of increased rates of natural disasters and variable weather patterns. One article reports that, "When a family is faced with the impact of the climate crisis, girls’ education is one of the first things families drop" (Medlicott, 2021). Another meta-analysis consisting of 53 studies report that two-thirds of those studies find that women are more often the victims of death or injury in the case of extreme weather events (Seller, 2016). If women are, on average, affected disproportionately affected by climate externalities, then one would expect that, on average, rational policymakers who are women would enact policies concerning climate policy at higher and potentially at more effective rates (pertaining to protecting the needs of women) relative to other compositions of policymakers with lower proportions of women, assuming policymakers who are women have the means and power to do so.

The outline of this paper will proceed as follows: I will first introduce the data I collected for this study and assess its reliability; I will then discuss the identification strategy I use to identify used to estimate the causal effect. I will then finish by discussing the results of my econometric analysis and its implications and then conclude.

# Data

The data used for this analysis was strategically gathered using publicly available sources (either by ethical web-scraping methods or direct download). The Inter-Parliamentary Union (IPU) maintains a consistent archive of records that monitors the number of women elected into the national legislature for each country. Not every government maintains the same structure of government, however, the IPU distinguishes between women elected into the lower-house and upper-house levels of government throughout multiple time periods. I take the average proportion of women holding office in the national legislature overall in a given year by taking a weighted average of the two houses. I obtained emissions data, including total yearly (gigatons of) $CO_2$ emissions per capita per country from the Emissions Database for Global Atmospheric Research (EDGAR) as maintained by the European Commission. I obtained country-year-level specific covariates, such as population and GDP per capita, from the World Bank. I obtained the data base linking each country's year they passed some legal right of suffrage from "A Lexical Index of Electoral Democracy" on the Harvard Dataverse (Skaaning et. al, 2015). I checked country-specific anomalies using data from the Pew Research Center (Schaeffer, 2020). Finally, I used U.S. Department of State's regional definitions to define which countries I define as either African or Arab.

By adjusting for time, country, and other specific endogenous effects on the proportion of women in government in a country ($c$) at year ($t$), I hope to show that, effectively, when proportions of women in government increase, $CO_2$ emissions decrease. However, one potential problem with data is that I fail to account for country-specific government intricacies that could prevent that could prevent women or a legislature from passing effective climate policy--such as failing to account for government corruption or other unique anomalies (wars between countries, coups, etc, that prevent a government from acting efficiently)--at every year. I assume that on average, these factors in the idiosyncratic term of my model are 0 for any given country at year $t$.

```{r visuals, echo=F, warning=F, message=F, fig.align='center', fig.width=16, fig.height=16}
label_data <- women %>%
  filter(year >= 1997) %>%
  group_by(countrycode) %>%
  summarize(
    ryear = sample(min(year):max(year), 1),
    womenrep = womenrep[which(year == ryear)],
    co2 = co2[which(year == ryear)]
  )

women %>% filter(year >= 1996) %>%
  ggplot(aes(x=year, y=womenrep, group=countrycode, color=countrycode))+
  geom_line()+
  geom_text_repel(
    data = label_data,
    aes(x=ryear, label = countrycode),
    label.padding = 0.25,
    box.padding = unit(0.5, "lines"),  # Padding around the box
    box.color = "black",  # Box color
    nudge_y = 0.1,  # Adjust position of labels
    direction = "both",
    max.overlaps = 20,
    segment.size = 0.4,
    nudge_x = 0.3,
    color="black"
  ) +
  labs(
    x="Year",
    y="Proportion of Women in Government",
    caption="Figure 1 - Average Proportion of Women in National Legislatures Over Time"
  )+
  guides(color = "none") +
  theme(axis.text = element_text(size = 36))+
  theme_minimal()

women %>% filter(year >= 1996) %>%
  ggplot(aes(x=year, y=co2, group=countrycode, color=countrycode))+
  geom_line()+
  geom_text_repel(
    data = label_data,
    aes(x=ryear, label = countrycode),
    label.padding = 0.25,
    box.padding = unit(0.5, "lines"),  # Padding around the box
    box.color = "black",  # Box color
    nudge_y = 0.3,  # Adjust position of labels
    direction = "both",
    segment.size = 0.4,
    nudge_x = 0.1,
    max.overlaps = 20,
    color="black"
  ) +
  guides(color = "none") +
  labs(
    x="Year",
    y=expression("Gigatons of "~CO[2]~" Emissions Per Capita"),
    caption="Figure 2 - Per Capita Carbon Emissions Over Time"
  )+
  theme(axis.text = element_text(size = 36))+
  theme_minimal()
```

In my sample, the United Arab Emirates (UAE) and Saudia Arabia are excluded in my sample because they don't give full suffrage to women: UAE held its first national elections in 2006, but only a select amount of men and women were eligible to participate (Schaeffer, 2020). Saudi Arabia doesn't hold national elections. Women are, however, enfranchised on the local level government. This leaves us with a total of 64 different countries in the sample. Using all the data which I was able to find data for the number of women in national legislature for a given country using data from the IPU, I was left with data spanning 1997-2022 when merged with $CO_2$ emissions data from EDGAR.

```{r summary_stats, echo=F}
women.subset <- women %>% filter(year >= 1997)

tibble(
  Variable = c("CO2 Emissions Per Capita", "Proportion of Women in Gvt", "GDP", "Population", "GDP per Capita", "Suffrage"),
  Mean = c(mean(women.subset$co2, na.rm = T) %>% round(3) %>% format(scientific=F), mean(women.subset$womenrep, na.rm=T) %>% round(3) %>% format(scientific=F), mean(women.subset$gdp, na.rm=T) %>% round(), mean(women.subset$population, na.rm=T) %>% round(3) %>% format(scientific=F),
           mean(women.subset$gdppercapita, na.rm=T) %>% round(3) %>% format(scientific=F),
           mean(women.subset$suffrage, na.rm=T) %>% round(3) %>% format(scientific=F)),
    SD = c(sd(women.subset$co2, na.rm = T) %>% round(3) %>% format(scientific=F), sd(women.subset$womenrep, na.rm=T) %>% round(3) %>% format(scientific=F),
           sd(women.subset$gdp, na.rm=T) %>% round(1), sd(women.subset$population, na.rm=T) %>% round(3) %>% format(scientific=F),
           sd(women.subset$gdppercapita, na.rm=T) %>% round(3) %>% format(scientific=F),
           sd(women.subset$suffrage, na.rm=T) %>% round(3) %>% format(scientific=F)),
    Min = c(min(women.subset$co2, na.rm = T) %>% round(3) %>% format(scientific=F), min(women.subset$womenrep, na.rm=T) %>% round(3) %>% format(scientific=F), min(women.subset$gdp, na.rm=T) %>% round(), min(women.subset$population, na.rm=T) %>% round(3) %>% format(scientific=F),
           min(women.subset$gdppercapita, na.rm=T) %>% round(3) %>% format(scientific=F),
           min(women.subset$suffrage, na.rm=T) %>% round(3) %>% format(scientific=F)),
    Max = c(max(women.subset$co2, na.rm = T) %>% round(3) %>% format(scientific=F), max(women.subset$womenrep, na.rm=T) %>% round(3) %>% format(scientific=F), max(women.subset$gdp, na.rm=T) %>% round(1), max(women.subset$population, na.rm=T) %>% round(3) %>% format(scientific=F),
           max(women.subset$gdppercapita, na.rm=T) %>% round(3) %>% format(scientific=F),
           max(women.subset$suffrage, na.rm=T) %>% round(3) %>% format(scientific=F)),
) %>% knitr::kable(caption = "Summary Statistics of Key Variables")

women.subset %>%
  group_by(countrycode) %>%
  summarize(
    N = length(!is.na(womenrep)),
    `CO2 Emissions Per Capita` = mean(co2, na.rm=T) %>% round(5),
    `Proportion of Women in Gvt` = mean(womenrep, na.rm=T) %>% round(5),
    GDP = mean(gdp, na.rm = T) %>% round(),
    Population = mean(population, na.rm=T) %>% round(3),
    `GDP per Capita` = mean(gdppercapita, na.rm=T) %>% round (5),
    Suffrage = first(suffrage)
  ) %>% knitr::kable(col.names = c("Country", "N", "CO2 Emissions Per Capita", "Proportion of Women in Gvt", "GDP", "Population", "GDP per Capita", "Suffrage"),
                     caption="Summary Statistics by Country") %>% kableExtra::kable_minimal() %>% kableExtra::kable_styling(font_size = 6)
```

# Identification

To estimate the causal effect that electing more women in national legislatures in African and Arab countries has on reducing per capita $CO_2$ emissions, I employ a two-stage difference-in-difference identification strategy.

Let $Y_{ct}$ represent the total (gigatons of) carbon emissions per capita for a country ($c$) in a year ($t$). I seek to identify consistently and interpret the causal effect of $\delta$, the effect of higher proportion of women in national legislatures in African and Arab countries on national $CO_2$ emissions per capita. Research from Mirziyoyeva & Salahodjaev (2023) suggests that women in government promote economic growth. This could pose potential problems of reverse causality: We need to be certain that any effect we see on women representation causes a reduction (or increase) in carbon emissions. If carbon emissions are highly correlated with economic growth and industrialization then an OLS regression on $Y_{ct}$ to estimate $\delta$ would almost certainly be biased. To first eliminate the potential for reverse causality--the supposition that perhaps lower carbon emissions causes a higher proportion of women to become elected during the same year; or rather, perhaps lower carbon emissions is correlated with other variables (such as economic and human development) that are linked to higher proportions of women becoming elected--I lag the treatment variable of interest, $W_{ct}$, the average proportion of women in national legislature for country $c$ during year $t$ one year. This implies that any change in $Y_{ct}$ will inherently come *after* a change in $W_{ct}$. Additionally, I introduce a matrix of country-year-specific covariates, $X_{ct}$, to adjust for additional confounding. This matrix includes a country's GDP per capita over time. This is also lagged one year to control for the fact that the effects on these covariates will necessarily imply that for any increase/decrease in $Y_{ct}$, the change will thus be associated with the previous year's economic growth for the respective country $c$.

In the two-stage design, I further suspect that there exists some endogeneity in $W_{ct}$ s.t.
$$
\exists \quad c_0, t_0 \quad s.t. \quad \mathbb{E}[W_{c_0t_0-1}\epsilon_{c_0t_0}]\neq0
$$
Both $CO_2$ emissions and women representation in government are highly correlated with industrialization and development. While $X_{ct}$ can eliminate some of the endogeneity of $W_{ct}$ by including GDP per capita, these covariates alone are not enough to explain a country's development. The rate at which a country develops, both industrially and politically, in regards to the countries in this study may be highly related to which European power originally occupied the territory and thus, how the power left the state when occupation was ceased, including the political institution(s) established. Endogeneity may also arise in what natural resources are available to the country at year $t$. Countries with a greater propensity to use carbon-intensive resources will thus have, on average, higher $CO_2$ emissions. If this also impacts women disproportionately such that women in these countries are more compelled to run for office and enact effective climate policies, then the estimand for $\delta$ will be biased in the negative direction.

To eliminate this endogeneity I invoke a two-stage difference-in-difference design, using an instrumental variable for $W_{ct-1}$. A common instrument in the literature for women parliament has been to use the number of years since the country passed a bill/law of suffrage (Mavisakalyan & Tarverdi, 2019). I use similar instrumental variable design here. Let $Z_{ct-1}$ be the number years since a country $c$ has granted suffrage (measured at year $t$, and lagged one year to align with $W_{ct}$).

In the difference-in-difference design, I use 62 country-fixed effects (omitting Angola and Burundi dummies due to colinearity). After lagging $W_{ct}$ one year, I am left with data spanning 1998-2022, thus, including year-fixed effects, I omit 1999. I utilize the following two-stage estimation approach:
$$
(1) \quad W_{ct-1}=\pi_0+\lambda Z_{ct-1}+X_{ct-1}'\Psi+\sum_{k=\text{Benin}}^{Zimbabwe}\mu_k\mathbb{1}(\text{Country}_c=k)+\sum_{j=1999}^{2022}\eta_j\mathbb{1}(\text{Year}_t=j)+v_{ct}
$$
$$
(2) \quad Y_{ct}=\beta_0+\delta W_{ct-1}+X_{ct-1}'\Omega+\sum_{k=\text{Benin}}^{Zimbabwe}\beta_k\mathbb{1}(\text{Country}_c=k)+\sum_{j=1999}^{2022}\gamma_j\mathbb{1}(\text{Year}_t=j)+\epsilon_{ct}
$$

## The validity of years since suffrage as an instrument

Mavisakalyan & Tarverdi, although make use of a cross-sectional data set, to their credit, report that, "an IV estimation can be potentially more convincing in a dataset that varies over time" (161). This is precisely what I employ here. I will expand on the research that uses years since suffrage as an instrument (Grier and Maldonado, 2015; Hicks et al., 2016). Relevance of $Z_{ct-1}$ is present with the idea that women in countries where suffrage is granted will see more involvement in government over time as women's rights expand in politics. Mavisakalyan & Tarverdi note that "a country’s history of suffrage should be highly relevant for female exposure to politics" (160).

```{r echo=F, message=F, warning=F, fig.width=8, fig.height=6, relevance}
women %>%
  ggplot(aes(y=lag(womenrep), x=lag(z)))+
  geom_point(size=0.5)+
  geom_jitter(size=0.5, width=0.05, height=0.05)+
  geom_smooth(method = "lm", se=F, color="#44CCAA")+
  theme_minimal()+
  labs(
    x=expression(Z[ct-1]),
    y=expression(W[ct-1]),
    caption = "Figure 3 - Relevance of Instrumental Variable"
  )

z.lm <- lm(lag(womenrep) ~ lag(z), data=women)
z.beta = coef(z.lm)[2]
z.se = summary(z.lm)$coefficients[2,2]
z.ci = (z.beta + c(-1,1)*z.se*qnorm(0.975)) %>% round(5)
```

The regression coefficient is significantly positive: `r z.beta %>% round(5)` (`r z.ci[1]`, `r z.ci[2]`).

With the difference-in-difference approach, using country and yearly fixed effects, there is reason to believe that $Y_{ct}(W_{ct-1}) \perp Z_{ct-1}|X_{ct-1}, \text{Country}_c, \text{Year}_t$. In other words, regardless of when a country granted suffrage, the potential resulting carbon emissions per capita as function of the proportion of women in the national legislature in a country $c$ are only dependent upon $W_{ct-1}$ after fixed effects are accounted for. However, as Mavisakalyan & Tarverdi note, "[our instrument] would be questionable if there is the possibility that the history of suffrage affects climate change policy outcomes through mechanisms other than the Female in Parliament." Mathematically, I assume,
$$
\mathbb{E}[Z_{ct-1}v_{ct}]=0\quad\forall c,t
$$
This could be violated if countries whose governments are established after colonial rule have (1) strong dependencies to the formal colonial power in terms of institutional structure and (2) as a function of that institutional structure, relies on carbon-intensive resources to kickstart its economy in post-colonial development. In other words, the instrument will be invalid if there are interactions between time, country, and carbon emissions. Evidently, more research needs to be done to see if this is the case, however, countries that granted suffrage to women immediately after occupation typically took place in the late 1900s, if it was less than random which countries these were, then the economic development--in relation to carbon-intensive resources--is also most likely non-random as well and thus would cause endogeneity in the instrument. In order to assert this, one would have to see a distinct pattern between the time periods of when women were granted suffrage and some other metric associated with carbon-intensive resource use such as oil exports. In this paper, I assume away this endogeneity by claiming that the instrument is as good as random (independent)--there is no distinct correlation between when a country granted suffrage and the types of resources the country typically consumes. Additionally, I assume monotonicity: the longer a country has passed from the year it granted suffrage, the more, on average, women are elected into national legislatures--I assume that there are no countries for which this trend reverses. The monotonicity assumption may be violated if women become less interested in politics over time for certain countries such that their participation decreases and hence the average proportion of women in national legislatures decrease with respect to years since suffrage.

## Assumptions for difference-in-differences

In the absence of more (or less) women being elected, the total $CO_2$ emissions per capita for each country would have followed the same trend over time as the other countries that did not elect more (or less) women in national legislatures at that specific time. Since my treatment variable ($W_{ct-1}$) is continuous, the parallel trends assumption is hard to verify. However, I assume no other events or factors are influencing $CO_2$ emissions per capita differently across different countries over time. In other words, I assume no unobserved heterogeneity; my estimate is robust to omitted covariates--this is most likely satisfied if we believe in the exclusion restriction of our instrumental variable.

Following Bertrand et. al (2004), I cluster standard errors by countries, where I assume standard errors across countries are independent from each other. This is most appropriate given that within each country, $CO_2$ emissions are most likely correlated with each other given that the economic and industrial structure mostly remains consistent within each country from year to year (yet different across country). This assumption could be violated if groups of countries form specific treaties and agreements to reduce $CO_2$ emissions or make other agreements pertaining to usage of other carbon-intensive resources.
```{r include_latex_table, fig.align='center', echo=F, warning=F, results='asis'}
cat(readLines("regression_results.tex"), sep = "\n")
```

# Results

The difference-in-difference parameter, $\delta$, was estimated using a two-stage design and assumed to be linear. If my assumptions hold, then the economic and statistic implications are relatively straightforward (interpreted on a 1% scale): On average, for every 1% the average proportion of women in national legislatures in African and Arab countries increase by, the total yearly (gigatons of) $CO_2$ emissions per capita decrease by 0.273 (significant at the 5% level). Interestingly, the effect on GDP per Capita (lagged one year), included in the $X_{ct-1}$ matrix, was found to have an insignificant effect, implying that the effects on GDP per capita growth (as a proxy variable for economic growth) on $CO_2$ per capita emissions can be almost entirely explained away by the year-level and country-level specific fixed effects.

## Economic implications for climate policy

Given that the assumptions hold, the fact that $\delta$ yields a significant effect implies that either (1) higher rates of being elected to office in national legislatures does, in fact, reduce yearly $CO_2$ emissions per capita by virtue empowering women to enact effective climate policies, or perhaps the more peculiar option, (2) the mere increased presence of women in national legislatures somehow either allows those with power to pass climate reform to pressures climate policy reform to take place. (1) Assumes that when women are elected into the national legislature is directly associated to a transfer of power to women to reform policies. We would hope this would be true, however, due to the high amount of corruption even in the high levels of government that has persisted throughout a subset of the countries in this study, one cannot conclude from this study alone that women in African and Arab nations are able to pass more effective climate policy reform (carbon emission reform in this case) when more women are present in the national legislature. Even though this hypothesis could absolutely be true as well, this study can neither confirm nor deny this. Thus, I will consider the possibility of (2): It could be that, coincidentally, when higher concentrations of women are present in national legislatures, the governments, as a whole, are more sensitive and cognizant of climate policy and not necessarily that women are the innate cause of the reduction in carbon emissions. Regardless of which hypothesis holds (or neither), it seems to be the case (with high confidence) that when more women enter the national legislature, we see lower per capita $CO_2$ emissions.

## Plausibility of Heterogeneous Treatment Effects

The estimate for $\delta$ is the local average treatment effect of increased proportions on reduced per capita $CO_2$ emissions for all countries. If the average proportion of women in national legislature affect per capita $CO_2$ emissions differently on a per country level, then each heterogeneous treatment effect would vary differently from $\delta$ (though centered around the true $\delta$). Realistically, heterogeneous effects are most likely present: Countries that have different dependencies on different carbon-intensive resources (such as for Qatar and Kuwait--see Figure 2) and countries that are more prone to corruption and war will have different priorities at the national government level. Thus, even as more women become elected, per capita $CO_2$ may not decrease depending on the current priorities of the country. A future study could examine the treatment effects for each country. This study assumes away heterogeneous treatment effects for the tradeoff of a larger sample size and a smaller standard error for $\hat{\delta}$.

# Conclusion

This paper reaffirms the effect that was outlined by Mavisakalyan & Tarverdi. Utilizing a causal two-stage difference-in-difference econometric framework, As national legislatures in African and Arab nations witness higher concentrations of women we see significantly lower per capita $CO_2$ emissions. While the exact mechanism through which women's presence influences climate policy remains open for further investigation, these findings hold significant policy implications. By encouraging women to run for office and dismantling barriers to their political participation, governments can potentially achieve environmental benefits alongside broader societal advancements. While this study focused on the average effect across countries, future research can delve deeper into heterogeneous treatment effects. Examining individual countries will allow for a more nuanced understanding of how factors like resource dependence and political structures interact with female representation to influence $CO_2$ emissions. Furthermore, investigating the specific policies enacted by legislatures with higher female representation can shed light on the mechanisms at play. This study underscores the importance of considering gender equality within climate change mitigation strategies. Understanding how women leadership shapes environmental policy is crucial for developing more effective strategies to achieve a sustainable future.

\newpage

# References

1. Emissions Database for Global Atmospheric Research (EDGAR). (n.d.). Edgar online. Retrieved from [https://edgar.jrc.ec.europa.eu/](https://edgar.jrc.ec.europa.eu/)

2. European Commission, Joint Research Centre (JRC). (2023). Emissions gap report 2023. Publications Office of the European Union. doi:10.2760/481442 [https://edgar.jrc.ec.europa.eu/report_2023](https://edgar.jrc.ec.europa.eu/report_2023)

3. Inter-Parliamentary Union (IPU). (n.d.). Women in national parliaments. [Data set]. Retrieved from [https://data.ipu.org/womon-ranking](https://data.ipu.org/womon-ranking)

4. Mavisakalyan, A., & Tarverdi, Y. (2019). Gender and climate change: Do female parliamentarians make a difference? Global Labor Organization (GLO) Discussion Paper No. 221. [https://www.sciencedirect.com/science/article/pii/S0176268017304500](https://www.sciencedirect.com/science/article/pii/S0176268017304500)

5. Medlicott, E. (2021, March 8). The five devastating reasons climate change affects women more than men. Retrieved from [https://www.euronews.com/green/2021/11/09/the-five-devastating-reasons-climate-change-affects-women-more-than-men](https://www.euronews.com/green/2021/11/09/the-five-devastating-reasons-climate-change-affects-women-more-than-men)

6. Mirziyoyeva, N., & Salahodjaev, D. (2023). Does representation of women in parliament promote economic growth? Considering evidence from Europe and Central Asia. Frontiers in Political Science, 4(2). [DOI: 10.3389/fpos.2023.1203221] [https://www.frontiersin.org/articles/10.3389/fpos.2023.1120287/full](https://www.frontiersin.org/articles/10.3389/fpos.2023.1120287/full)

7. Schaeffer, Katherine (2020, October 5). Pew Research Center. Key facts about women's suffrage around the world, a century after U.S. ratified 19th Amendment. Retrieved April 1, 2024, from [https://www.pewresearch.org/short-reads/2020/10/05/key-facts-about-womens-suffrage-around-the-world-a-century-after-u-s-ratified-19th-amendment/](https://www.pewresearch.org/short-reads/2020/10/05/key-facts-about-womens-suffrage-around-the-world-a-century-after-u-s-ratified-19th-amendment/)

8. Sellers, S. (2016). Gender and climate change: A closer look at existing evidence. Global Gender and Climate Alliance (GGCA). [https://wedo.org/gender-and-climate-change-a-closer-look-at-existing-evidence-ggca/](https://wedo.org/gender-and-climate-change-a-closer-look-at-existing-evidence-ggca/)

9. Skaaning, Svend-Erik; John Gerring; Henrikas Bartusevicius, 2015, "A Lexicial Index of Electoral Democracy", [https://doi.org/10.7910/DVN/29106](https://doi.org/10.7910/DVN/29106), Harvard Dataverse, V6

10. United Nations Development Programme (UNDP). (2022, March 8). International Women’s Day 2022 “Gender equality today for a sustainable tomorrow”. Retrieved from [https://www.undp.org/speeches/undp-administrators-statement-international-womens-day-2023-8-march](https://www.undp.org/speeches/undp-administrators-statement-international-womens-day-2023-8-march)

11. U.S. Department of State. (n.d.). Countries and Areas List. Retrieved from [https://www.state.gov/countries-and-areas-list/](https://www.state.gov/countries-and-areas-list/)

12. The World Bank. (n.d.). World Bank Open Data. Retrieved from [https://databank.worldbank.org](https://databank.worldbank.org)

13. Grier, R., & Maldonado, B. (2015). Electoral experience, institutional quality, and economic development in Latin America. Oxford Development Studies, 43(2), 253-280.

14. Hicks, D. L., Hicks, J. H., & Maldonado, B. (2016). Women as policy makers and donors: Female legislators and foreign aid. European Journal of Political Economy, 44(2), 118-135. doi: 10.1016/j.europoleco.2015.12.004