---
title: "Cluster-Robust Variance Estimators for Instrumental Variables"
subtitle: "Theory and Application in Assessing the Impact of Gender Representation on Carbon Emissions"
author: "Sam Lee"
format: pdf
header-includes:
  - \usepackage{fontspec}
  - \setmainfont{Times New Roman}
  - \usepackage{setspace}
  - \renewcommand{\normalsize}{\fontsize{12}{14.4}\selectfont}
  - \usepackage[margin=1in]{geometry}
---

```{r setup, include=FALSE}
#Additional packages used in this coument
library(patchwork)
library(stargazer)
library(Matrix)
library(expm)
library(ivreg)
library(sandwich)

set.seed(04142024)

#Load in the data matrices for the applied example
source("CRVE/setup.R")
#Load in cluster robust formula
source("CRVE/cluster.robust.R")
#Load in function for data-generating bootstrap process
source("CRVE/DGP.R")
#Load in function to calculate 2sls regression
source("CRVE/regression.R")
#Function to run simulations
source("CRVE/simulate.R")
```
# Abstract
```{=latex}
\begin{doublespace}
```
Within applied econometrics, the practice of clustering standard errors by groups that are presumed to have errors correlated within the group but otherwise uncorrelated between other groups has invoked a series of large research on econometric methodology, especially within the past couple of decades. One recent paper in the literature outlines a guide to empirical practice using cluster-robust variance estimators (CRVEs) (MacKinnon et. al 2022). The authors of this paper assess three feasible CRVEs, with particular emphasis on a jackknife estimator. While the authors of this paper assert "for such models [with clustered data estimated by instrumental variables (IV)], neither the current state of econometric theory nor the available simulation evidence allows [the authors] to make recommendations with any confidence," this paper attempts to expand on the guide on empirical practice made by MacKinnon et. al and apply theory to an empirical analysis using instrumental variables. This paper examines what the effect of electing higher proportions of women in African and Arab nations has on yearly $CO_2$ emissions per capita using years since suffrage as an instrumental variable. To assess the validity of inference, I first propose three feasible cluster-robust variance estimators for a GMM framework in the just-identified case following MacKinnon et. al. I apply the proposed estimators to an empirical example to assess the credibility of inference and to evaluate their long-run performance using a Monte Carlo analysis. I find that one particular estimator may possibly perform more consistently than the others, while the others tend to under reject. 
```{=latex}
\end{doublespace}
```
# I. Introduction
```{=latex}
\begin{doublespace}
```
The most well-known CRVE used in econometric practice is by far the Liang-Zeger estimator (1986). This formula relies on the asymptotic properties of the OLS estimator. Furthermore, one novel critique of the difference-in-differences (DiD) estimator argues that applied econometric approaches using a DiD approach will typically underestimate the standard deviation of treatment effect without using a proper CRVE (Bertrand et. al 2004). Notably, the paper from Bertand et. al uses a Monte Carlo analysis to generate new random vectors for the response variables to illustrate the empirical coverage of CRVE estimators for both a large and small amount of clusters. In this paper, I will first introduce the CRVEs that MacKinnon et. al propose as standard practice in applied work. Critically, since the CRVEs used in mainstream practice rely on the asymptotic properties of the OLS estimator, I attempt to parallel the work to a generalized method of moments (GMM) framework suitable for instrumental variables. The objective of this paper is to assess the performance of these proposed estimators and the implications on an important issue in society. In this paper, I analyze the effect of how electing higher proportions of women in African and Arab countries can decrease yearly per capita $CO_2$ emissions. I discuss why this is an important question both socially and econometrically. I will then explain the data-generating process I will use for my Monte Carlo analysis, my results, and then conclude.
```{=latex}
\end{doublespace}
```
# II. Three Feasible CRVEs
```{=latex}
\begin{doublespace}
```
The econometric guide from MacKinnon et. al discusses in detail three feasible CRVEs for the following linear regression model:
```{=latex}
\end{doublespace}
```
$$
(1)\quad y_i=x_i'\beta+u_i
$$
```{=latex}
\begin{doublespace}
```
In practice, if the data is divided up into $G$ \# of disjoint clusters. Equation (1) can thus be equivalently written as,
```{=latex}
\end{doublespace}
```
$$
(2)\quad y_g=X_g\beta+u_g,\quad g=1,...,G
$$
```{=latex}
\begin{doublespace}
```
While this paper serves a purpose to not reinvent the wheel, and rather, works to provide enough preliminary detail so that readers with sufficient econometric backgrounds may understand the steps I took to arrive at my own CRVE, it helps to see the groundwork that previous econometricians have laid before. That being said, without going into too much detail that MacKinnon et. al (among so many others) provide, it follows that the asymptotic properties of (1) follow the usual OLS properties. Assuming data are generated at $\beta=\beta_0$:
```{=latex}
\end{doublespace}
```
$$
\sqrt{n}(\hat{b}-\beta_0)\underset{d}{\rightarrow}N(0,(\mathbb{E}[X_iX_i']^{-1}\mathbb{E}[X_iX_i'u_i^2]\mathbb{E}[X_iX_i']^{-1})
$$
```{=latex}
\begin{doublespace}
```
Since $X$ is merely the stacked matrix of the clusters, the only difference between the asymptotic properties of $\hat{b}$ in (1) and (2) is the calculation of $\mathbb{E}[x_ix_i'u_i^2]$. Empirically, for (1), this can be estimated with the sample mean (robust only to unknown forms of heteroskedasticity):
```{=latex}
\end{doublespace}
```
$$
\frac{1}{n}\sum_{i=1}^nX_iX_i'\hat{u}_i^2
$$
```{=latex}
\begin{doublespace}
```
Thus, assuming data are divided into $G$ \# of disjoint groups, a convenient way to cluster standard errors by each group has been to rewrite the summation above as disjoint multiplicative matrix additions, summing the residual score vectors across each group $g$. Hence, if there is complete homoskedasticity, it follows that,
```{=latex}
\end{doublespace}
```
$$
\frac{1}{n}\sum_{i=1}^nX_iX_i'\hat{u}_i^2=\frac{1}{n}\sum_{g=1}^GX_g'\hat{u}_g\hat{u}_g'X_g
$$
```{=latex}
\begin{doublespace}
```
Of course, the equality won't hold if there is correlation *within* each cluster, hence the need to use an estimator for a cluster-robust variance. Thus, the first CRVE[^20] MacKinnon provide follows directly as,

[^20]: Note that this is the variance of the random variable, $\sqrt{n}(\hat{b}-\beta_0)$.
```{=latex}
\end{doublespace}
```
$$
\text{CV}_1: n\frac{G(N-1)}{(G-1)(N-k)}(X'X)^{-1}(\sum_{g=1}^G\hat{s}_g\hat{s}_g')(X'X)^{-1};\quad\hat{s}_g=X_g'\hat{u}_g
$$
```{=latex}
\begin{doublespace}
```
Where a correction for degrees of freedom is applied. While MacKinnon, et. al note that this is the most widely used cluster-robust formula in use, the authors keenly mention that $X_g'\hat{u}_g$ is not always a good estimator for $s_g$. Two alternative CRVEs[^1] have been proposed by the literature to address this issue by transforming the residual score vectors ($\hat{s}_g$) (Bell & McCaffrey, 2002).

[^1]: As MacKinnon et. al pay particular interest to $\text{CV}_3$. As the authors note, this cluster-robust variance estimator is a jackknife estimator and is discussed thoroughly throughout their paper.

```{=latex}
\end{doublespace}
```
$$
\text{CV}_2: n(X'X)^{-1}(\sum_{g=1}^{G}\grave{s}_g\grave{s}_g')(X'X)^{-1}; \quad \grave{s}_g=X'M_{gg}^{-1/2}\hat{u}_g;\quad M_{gg}=I_{N_g}-X_g(X'X)^{-1}X_g'
$$
$$
\text{CV}_3: n\frac{G-1}{G}(X'X)^{-1}(\sum_{g=1}^{G}\acute{s}_g\acute{s}_g')(X'X)^{-1}; \quad \acute{s}_g=X'M_{gg}^{-1}\hat{u}_g
$$

# III. A Cluster-Robust Variance Estimator for Instrumental Variables
```{=latex}
\begin{doublespace}
```
While the empirical guide outlined by MacKinnon et. al is useful and practical in its own right, it has its limitations. The cluster-robust variance estimators discussed in the paper only make use of the empirical matrix $X$, an $N \times k$ matrix presumed to have none of its independent variables violating endogeneity. The authors candidly assert that "for such models [with clustered data estimated by instrumental variables (IV)], neither the current state of econometric theory nor the available simulation evidence allows [the authors] to make recommendations with any confidence." Thus, as a matter of econometric exploration, I attempt to push this boundary by using the pre-existing econometric research and literature to propose a cluster-robust variance estimator for instrumental variables in the generalized method of moments (GMM). For the simplification of this paper, I only consider the just-identified case.
```{=latex}
\end{doublespace}
```
## GMM Framework for Instrumental Variables
```{=latex}
\begin{doublespace}
```
In the GMM framework, we will first consider the moment condition for instrumental variables:
```{=latex}
\end{doublespace}
```
$$
\mathbb{E}[m_i(\beta)=Z_i(Y_i-X_i'\beta)]=0; \quad G(\beta) := \left[ \frac{\partial m_i(\beta)}{\partial\beta'} \right]=-\mathbb{E}[Z_iX_i'] \quad (\text{has rank } k)
$$
```{=latex}
\begin{doublespace}
```
Assuming conditions for identification are met[^2], for data generated with $\beta=\beta_0$, the asymptotic properties of GMM estimation follow:

[^2]: While I don't discuss the assumptions and conditions for identification in detail in this paper, I do discuss in detail the assumptions for IV identification in my applied example in another paper [here](https://github.com/SamLeeBYU/Elections/blob/main/Report.pdf). However, in addition to the assumptions briefly mentioned above, I also assume that $\beta$ to uniquely solves the moment condition.

```{=latex}
\end{doublespace}
```
$$
\sqrt{n}(\hat{b}-\beta_0)\underset{d}{\rightarrow}N(0,\left[G(\beta)\mathbb{E}[m_i(\beta)m_i(\beta)']^{-1}G(\beta)' \right]^{-1})\implies
$$
$$
\sqrt{n}(\hat{b}-\beta_0)\underset{d}{\rightarrow}N(0, \left[\mathbb{E}[(X_iZ_i)]\mathbb{E}[Z_iZ_i'u_i^2]^{-1}\mathbb{E}[(X_iZ_i)' \right]^{-1})
$$
```{=latex}
\begin{doublespace}
```
The objective of this section will be to outline a reasonable cluster-robust variance estimator following the research that has been done on this topic. Moving from equation (1) to a two-stage least squares design, assume data is generated from the following just-identified two-stage least squares process, where $w_i$ is endogenous. Assume exclusion, independence, and relevance conditions are all met for the instrument $z_i$. Let the empirical just-identified two-stage regression be characterized as
```{=latex}
\end{doublespace}
```
$$
(3)\quad w_i=z_i'\lambda+v_i
$$
$$
(4)\quad y_i=\left[ w_i | q_i \right]'\left[ \delta\over\alpha \right] + \epsilon_i = x_i'\beta+\epsilon_i
$$
```{=latex}
\begin{doublespace}
```
Where $\delta$ is the (local average) treatment effect of interest on $w_i$. Let $q_i$ be a vector of included covariates[^3] with corresponding effects, $\alpha$.

[^3]: I notate equation (4) with augmented vectors to (i) show how I structure the data in the code and (ii) to show the connection between the first and second stages.

```{=latex}
\end{doublespace}
```
```{=latex}
\begin{doublespace}
```
Thus, let the variance estimator of interest from the GMM framework be defined as,
```{=latex}
\end{doublespace}
```
$$
V = \left[\mathbb{E}[(X_iZ_i)]\mathbb{E}[Z_iZ_i'u_i^2]^{-1}\mathbb{E}[(X_iZ_i)' \right]^{-1}
$$
```{=latex}
\begin{doublespace}
```
Using the sample means as empirical estimates, an estimate for $V$ reasonably follows as[^4] (assuming homoskedasticity),

[^4]: By the Law of Large Numbers and the Continuous Mapping Theorem, $\hat{V}$ converges to $V$.
```{=latex}
\end{doublespace}
```
$$
\hat{V}= \left[\left(\frac{1}{n}X'Z\right)\left(\frac{1}{n}Z'Z\hat{u}^2\right)^{-1}\left(\frac{1}{n}X'Z\right)'\right]^{-1}
$$
```{=latex}
\begin{doublespace}
```
Now, assuming the data can be classified into $G$ \# of disjoint clusters, using the same logic that by which the OLS CRVE was derived, $\hat{V}$ can be respectively rewritten as:
```{=latex}
\end{doublespace}
```
$$
\hat{V}_g= \left[\left(\frac{1}{n}X'Z\right)\left(\frac{1}{n}\sum_{g=1}^GZ_g'\hat{u}_g\hat{u}_g'Z_g \right)^{-1}\left(\frac{1}{n}X'Z\right)'\right]^{-1}
$$
```{=latex}
\begin{doublespace}
```
Simplifying, we arrive at our first CRVE for instrumental variables in a GMM framework (and by including the degrees of freedom correction[^0] from $\text{CV}_1$):

[^0]: Interestingly enough, Stata uses this CRVE when the `cluster` option is specified in its `ivregress` command. However, its formula ignores the degrees of freedom correction.

```{=latex}
\end{doublespace}
```
$$
\overset{iv}{CV}_{1}: n\frac{G(N-1)}{(G-1)(N-k)}\left[\left(X'Z\right)\left(\sum_{g=1}^G\hat{\zeta}_g\hat{\zeta}_g' \right)^{-1}\left(X'Z\right)'\right]^{-1}; \quad \hat{\zeta}_g=Z'_g\hat{u_g}
$$
```{=latex}
\begin{doublespace}
```
Thus, the OLS cluster-robust variance estimators from Section II of this paper can all be rewritten along the same line of theory:
```{=latex}
\end{doublespace}
```
$$
\overset{iv}{CV}_{2}: n\left[\left(X'Z\right)\left(\sum_{g=1}^G\grave{\zeta}_g\grave{\zeta}_g' \right)^{-1}\left(X'Z\right)'\right]^{-1}; \quad \grave{\zeta}_g=Z'_g\overset{iv}{M}_{gg}^{-1/2}\hat{u_g}; \quad \overset{iv}{M}_{gg}=I_{N_g}-Z_g(Z'Z)^{-1}Z_g'
$$
$$
\overset{iv}{CV}_{3}: \frac{n(G-1)}{G}\left[\left(X'Z\right)\left(\sum_{g=1}^G\acute{\zeta}_g\acute{\zeta}_g' \right)^{-1}\left(X'Z\right)'\right]^{-1}; \quad \acute{\zeta}_g=Z'_g\overset{iv}{M}_{gg}^{-1}\hat{u_g}
$$
```{=latex}
\begin{doublespace}
```
The goal of this paper will be to show whether the asymptotic properties of these estimators hold (if any). To demonstrate this, I will now apply these formulas to an applied example that uses instrumental variables.
```{=latex}
\end{doublespace}
```
# IV. Empirical Example

## Do Increased Proportions of Women in National Legislatures of African and Arab Countries Significantly Decrease per Capita $CO_2$ Emissions?
```{=latex}
\begin{doublespace}
```
I seek to identify the causal effect that electing higher proportions of
women in national legislatures in African and Arab nations has on yearly
per capita $CO_2$ emissions. I utilize a two-stage
difference-in-difference approach on 64 different countries throughout
Africa and the Middle East from 1998-2022. To eliminate the potential
endogeneity in using the proportion of women in national legislatures as
a treatment on per capita $CO_2$ emissions, I use years since women were
granted suffrage in each respective country as an instrumental variable.
One nominal paper in the literature has shown evidence that increased
proportions of women in parliaments are more than likely causally related
to stricter climate policies (Mavisakalyan & Tarverdi, 2019). However,
this paper does not take into account how carbon emissions and women in
government have evolved. To my knowledge, little research has
been done to show an intertemporal causal link between women in
parliaments and its effects on climate. I expand on this research by
showing how carbon emissions have changed over time to elicit a local
average treatment effect between changes in gender compositions at the
national legislature level.[^5].

[^5]: The full applied analysis for this project can be
found on this project's Github repository: [https://github.com/SamLeeBYU/Elections/Report.pdf](https://github.com/SamLeeBYU/Elections/Report.pdf).
```{=latex}
\end{doublespace}
```
## Background
```{=latex}
\begin{doublespace}
```
The current climate and economic research today overwhelmingly support
the thesis that women---especially in developing countries---are
disproportionately affected by the negative externalities of climate
change. During periods of increased droughts, which are exacerbated due to the
effects of climate change, women often make long trips to meet
agricultural, hygiene, and family needs. This has been linked to
increased violence against women (UNDP, 2019). Increased risks of
poverty and food insecurities have also been shown to increase as a result
of increased rates of natural disasters and variable weather patterns.
One article reports that "when a family is faced with the impact of the
climate crisis, girls' education is one of the first things families
drop" (Medlicott, 2021). Another meta-analysis consisting of 53 studies
report that two-thirds of those studies find that women are more often
the victims of death or injury in the case of extreme weather events
(Seller, 2016). 

If women are disproportionately affected by climate
externalities, one would expect that female policymakers would enact
effective climate policy at higher rates relative to other policymakers,
assuming they have the means and power to do so. Thus, the causal
question of interest reasonably follows as to whether increased proportions
of women decrease $CO_2$ emissions. As we will see, the nature of
causality of this question is rather difficult to elicit due to
intricate endogeneities that persist throughout this problem. 

The purpose of the empirical section of this paper seeks not to convince readers
that the complex problem of endogeneity and causality can be resolved (although these issues are intricately discussed in my [applied work](https://github.com/SamLeeBYU/Elections/Report.pdf)). Rather, this empirical example is included for two motivating reasons: (1) The
necessity and obligation of answering the question at hand as a result of expanding on Mavisakalyan and Tarverdi's research has great implications for climate policy, and (2) The econometric implications of using a properly adjusted cluster-robust variance estimator have similar impacts for the reliability of conclusions derived from this study and the econometrics field as a whole.

```{=latex}
\end{doublespace}
```
## Data
```{=latex}
\begin{doublespace}
```
The data used for this analysis was strategically gathered using
publicly available sources[^6]. The Inter-Parliamentary Union (IPU)
maintains a consistent archive of records that monitors the number of
women elected into the national legislature for each country.
Not every government maintains the same structure of government, however, the IPU
distinguishes between women elected into the lower house and upper house
levels of government throughout multiple periods. I take the
average proportion of women holding office in the national legislature
overall in a given year by taking a weighted average of the two
houses[^7] (See Figure 1). I obtained emissions data, including total yearly
(gigatons of) $CO_2$ emissions per capita per country, from the
Emissions Database for Global Atmospheric Research (EDGAR) as maintained
by the European Commission (See Figure 2).

[^6]: Data was collected using ethical web-scraping methods or obtained
    via direct download. Further documentation and the scripts used to
    obtain the data used for this analysis can be found on this
    project's Github repository
    <https://github.com/SamLeeBYU/Elections>.
[^7]: Econometrically, this assumes that on average, the power that women are given in each house of national legislature is equal to the 'weight' of women in that chamber of legislature: In other words, assuming a fair political system, I assume no chamber can systemically dominate the other chamber.

```{r visuals, echo=F, warning=F, message=F, fig.align='center', fig.width=16, fig.height=16}
women <- read_dta("Data/womenrep.dta")

women %>% filter(year >= 1996) %>%
  ggplot(aes(x=year, y=womenrep, group=country))+
  geom_line()+
  labs(
    x="Year",
    y="Proportion of Women in Government",
    title="Figure 1 - Average Proportion of Women in National Legislatures Over Time"
  )+
  guides(color = "none") +
  theme(axis.text = element_text(size = 36))+
  facet_wrap(.~country, nrow=8)+
  scale_x_continuous(labels = function(x) substr(x, 3, 4))+
  theme_minimal()

women %>% filter(year >= 1996) %>%
  ggplot(aes(x=year, y=co2, group=country))+
  geom_line()+
  guides(color = "none") +
  labs(
    x="Year",
    y=expression("Gigatons of "~CO[2]~" Emissions Per Capita"),
    title="Figure 2 - Per Capita Carbon Emissions Over Time"
  )+
  facet_wrap(.~country, nrow=8, scales = "free_y")+
  scale_x_continuous(labels = function(x) substr(x, 3, 4))+
  theme(axis.text = element_text(size = 36))+
  theme_minimal()
```

I obtained country-year-level specific
covariates, such as population and GDP, from the World Bank (See Table 1 for summary statistics on key variables). I
obtained the database linking each country's year they passed a right
of universal suffrage from "A Lexical Index of Electoral Democracy" on
the Harvard Dataverse (Skaaning et. al, 2015). I checked
country-specific anomalies using research from the Pew Research
Center (Schaeffer, 2020).  Not all countries in Africa and the Middle East treat women equally. Even when suffrage is granted, it is not granted universally. Thus, the instrument may not influence the proportion of women in government the same in every country. This does not discredit my econometric analysis, however, so long as the monotonicity assumption of my instrument is met. However, I exclude obvious anomalies in my analytic sample (See Table 2 for summary statistics for each country). In my sample, the United Arab Emirates (UAE) and Saudia Arabia are excluded because they don't give full suffrage to women[^8]. Finally, I used the U.S. Department of State's
regional definitions to define which countries I define as either African or Arab.

[^8]: The UAE held
its first national elections in 2006, but only a select amount of men and women were eligible to participate (Schaeffer, 2020). Saudi Arabia doesn't hold national elections. Women are, however, enfranchised at the local level government levels.
```{=latex}
\end{doublespace}
```
```{r summary_stats, echo=F}
women <- read_dta("Data/womenrep.dta")

women.subset <- women %>% filter(year >= 1997)

tibble(
  Variable = c("CO2 Emissions Per Capita", "Proportion of Women in Gvt", "ln(GDP)", "ln(Population)", "Suffrage"),
  Mean = c(mean(women.subset$co2, na.rm = T) %>% round(3) %>% format(scientific=F), mean(women.subset$womenrep, na.rm=T) %>% round(3) %>% format(scientific=F), mean(log(women.subset$gdp), na.rm=T) %>% round(), mean(log(women.subset$population), na.rm=T) %>% round(3) %>% format(scientific=F),
           mean(women.subset$suffrage, na.rm=T) %>% round(3) %>% format(scientific=F)),
    SD = c(sd(women.subset$co2, na.rm = T) %>% round(3) %>% format(scientific=F), sd(women.subset$womenrep, na.rm=T) %>% round(3) %>% format(scientific=F),
           sd(log(women.subset$gdp), na.rm=T) %>% round(1), sd(log(women.subset$population), na.rm=T) %>% round(3) %>% format(scientific=F),
           sd(women.subset$suffrage, na.rm=T) %>% round(3) %>% format(scientific=F)),
    Min = c(min(women.subset$co2, na.rm = T) %>% round(3) %>% format(scientific=F), min(women.subset$womenrep, na.rm=T) %>% round(3) %>% format(scientific=F), min(log(women.subset$gdp), na.rm=T) %>% round(), min(log(women.subset$population), na.rm=T) %>% round(3) %>% format(scientific=F),
           min(women.subset$suffrage, na.rm=T) %>% round(3) %>% format(scientific=F)),
    Max = c(max(women.subset$co2, na.rm = T) %>% round(3) %>% format(scientific=F), max(women.subset$womenrep, na.rm=T) %>% round(3) %>% format(scientific=F), max(log(women.subset$gdp), na.rm=T) %>% round(1), max(log(women.subset$population), na.rm=T) %>% round(3) %>% format(scientific=F),
           max(women.subset$suffrage, na.rm=T) %>% round(3) %>% format(scientific=F)),
) %>% knitr::kable(caption = "Summary Statistics of Key Variables")
```
```{=latex}
\begin{doublespace}
```
One potential problem with the
data is that I do not account for country-specific government
intricacies that could prevent women or a legislature from
passing effective climate policy---such as not accounting for government
corruption or other unique anomalies (wars between countries, coups,
laws that are discriminatory against women, etc, that prevent a
government from acting efficiently)---at every year[^9]. Merging the suffrage data with all the IPU available data and the $CO_2$ emissions data from EDGAR, I was left with 64 countries spanning 1997-2022 [^10].

[^9]: To assert consistency in my estimates, econometrically, I assume
    that on average, these factors in the idiosyncratic term of my
    econometric model are 0 for any given country at year $t$. (See
    equation (7)).
    
[^10]: Notably, the data is left as an unbalanced panel data set (Refer to Table 2). The World Bank does not have GDP or population data for every single year of every single country in the data, and critically, the IPU also notes that it cannot reliably retrieve the number of women in national legislatures for every year. These years for those specific countries are naturally excluded. I assume that the probability at which data for a specific year-country combination will be missing is essentially random. This assumption may be violated if countries that are systemically difficult to obtain data for (such as Somalia or the Central African Republic) is also correlated with other covariates and or $CO_2$ emissions.

```{=latex}
\end{doublespace}
```
```{r country_stats, echo=F}
women.subset %>%
  group_by(country) %>%
  summarize(
    N = sum(!is.na(womenrep)),
    `CO2 Emissions Per Capita` = mean(co2, na.rm=T) %>% round(5),
    `Proportion of Women in Gvt` = mean(womenrep, na.rm=T) %>% round(5),
    GDP = mean(gdp, na.rm = T) %>% round(),
    Population = mean(population, na.rm=T) %>% round(3),
    Suffrage = first(suffrage)
  ) %>% knitr::kable(col.names = c("Country", "N", "Mean CO2 Emissions Per Capita", "Mean Proportion of Women in Gvt", "Mean GDP", "Mean Population", "Suffrage"),
                     caption="Summary Statistics by Country") %>% kableExtra::kable_minimal() %>% kableExtra::kable_styling(font_size = 6)
```

## Identification

```{=latex}
\begin{doublespace}
```
To estimate the causal effect that electing more women in national
legislatures in African and Arab countries has on reducing per capita
$CO_2$ emissions, I employ a two-stage difference-in-difference
identification strategy. In regard to curbing the negative
externalities of climate change in the context of this paper, I use per
capita $CO_2$ emissions as a response variable. Let $Y_{ct}$ represent
the total (gigatons of) carbon emissions per capita for a country ($c$)
in a year ($t$). I seek to identify consistently and interpret the local
average treatment effect of $\delta$, the effect of a higher proportion of
women in national legislatures in African and Arab countries on national
$CO_2$ emissions per capita. Research from Mirziyoyeva and Salahodjaev
(2023) suggests that women in government promote economic growth. This
could pose potential problems of reverse causality: We need to be
certain that any effect we see on women's representation causes a
reduction (or increase) in carbon emissions. If carbon emissions are
highly correlated with economic growth and industrialization then an OLS
regression (that is, using a typical difference-in-differences approach)
on $Y_{ct}$ to estimate $\delta$ would almost certainly be biased and
inconsistent. To first eliminate the endogeneity for reverse
causality---the supposition that perhaps lower carbon emissions cause a
higher proportion of women to become elected during the same year; or
rather, perhaps higher carbon emissions are correlated with other
variables such as economic growth that are linked to higher proportions
of women becoming elected---I lag the treatment variable of interest,
$W_{ct}$, the average proportion of women in national legislature for
country $c$ during year $t$ one year. This implies that any change in
$Y_{ct}$ will inherently come *after* a change in $W_{ct}$. In the two-stage design, I further suspect that there exists some
endogeneity in $W_{ct}$[^11]. Both $CO_2$ emissions and women's
representation in government are highly correlated with
industrialization and development. While I do
not focus on specific regulatory practices within this paper, an a priori assumption could be that effective
climate policy would consist of enacting particular policies and
regulations that would reduce the $CO_2$ emissions by economic
constraint. It could be that increased proportions of women in government could
(causally) affect economic performance within a country at time $t$. Hence, I include a sequence of matrices of autoregressive economic controls, represented by $\sum_{p=2}^PX_{ct-p}$ (where $p$ is the $p$th lag of the economic covariate[^12]).
Thus, to establish the causal chain of inference, I infer:
```{=latex}
\end{doublespace}
```
```{=latex}
\begin{doublespace}
```
$$
\sum_{p=2}^PX_{ct-p}\rightarrow W_{ct-1}\rightarrow Y_{ct}
$$
Where $X_{ct-p}$ is a matrix of controls related to economic growth.
For the same reason the $W_{ct}$ is lagged one year to ensure that a
change in $Y_{ct}$ comes after a change in $W_{ct}$, I include a series
of autoregressive lags in the economic controls (starting at lag 2) to
ensure that changes in $W_{ct-1}$ occur *after* changes in economic
development. For this two-stage regression, $X_{ct-p}$ consists of logged GDP and logged population for
a country at year $t$[^13]. While $X_{ct}$ can eliminate some of
the endogeneity of $W_{ct}$ by including economic controls, these
covariates alone are not enough to explain a country's development and
how it interacts with how women are elected into national legislatures.
The rate at which a country develops, both industrially and politically,
in regards to the countries in this study may be highly related to which
European power originally occupied the territory and thus, how the power
left the state when the occupation ceased, including the political
institution(s) established. Endogeneity may also arise in what natural
resources are available to the country at year $t$. Countries with a
greater propensity to use carbon-intensive resources will thus have, on
average, higher $CO_2$ emissions. If this also impacts women
disproportionately such that women in these countries are more compelled
to run for office and enact effective climate policies, then the
estimand for $\delta$ will be biased in the negative direction.

[^11]: Mathematically, with endogeneity in the treatment variable, I
    assume
    $\exists \quad c_0, t_0 \quad s.t. \quad \mathbb{E}[W_{c_0t_0-1}\epsilon_{c_0t_0}]\neq0$.
[^12]: A priori I set $P=5$, the total number of lags on all the economic covariate controls.
[^13]: A further analysis could (and should) include further economic
controls that may affect $W_{ct-1}$ such as fertility rate, etc.

To eliminate this endogeneity I invoke a two-stage
difference-in-difference design, using an instrumental variable for
$W_{ct-1}$. A common instrument in the literature for women's representation in parliament
has been to use the number of years since the country granted suffrage
(Mavisakalyan & Tarverdi, 2019). I use a similar instrumental variable
design here. Let $Z_{ct-1}$ be the number of years since a country $c$ has
granted suffrage (measured at year $t$, and lagged one year to align
with $W_{ct-1}$). In the difference-in-difference design, I use 62 country-fixed effects
(omitting the Qatar and Iraq dummies due to colinearity). After lagging
$W_{ct}$ one year, I am left with data spanning 1998-2022, thus,
including year-fixed effects, I also omit 1998 for colinearity. I
utilize the following two-stage estimation approach:
```{=latex}
\end{doublespace}
```
$$
(5) \quad W_{ct-1}=\pi_0+\lambda Z_{ct-1}+\sum_{p=2}^PX_{ct-p}'\Psi_p+
$$
$$
\sum_{k=\text{Angola }k\neq\text{Qatar, Iraq}}^{Zimbabwe}\mu_k\mathbb{1}(\text{Country}_c=k)+\sum_{j=1999}^{2022}\eta_j\mathbb{1}(\text{Year}_t=j)+v_{ct}
$$
$$
(6) \quad Y_{ct}=\beta_0+\delta W_{ct-1}+\sum_{p=2}^PX_{ct-p}'\Omega_p+
$$
$$
\sum_{k=\text{Angola; }k\neq\text{Qatar, Iraq}}^{Zimbabwe}\beta_k\mathbb{1}(\text{Country}_c=k)+\sum_{j=1999}^{2022}\gamma_j\mathbb{1}(\text{Year}_t=j)+\epsilon_{ct}
$$
```{=latex}
\begin{doublespace}
```
It is worth noting that equations (5) and (6) are of the two-stage form found in equations (3) and (4), though with a lot more added covariates. More specifically, the data in this example can divided into $G$ \# of disjoint groups. Clusters in this example could be defined as either years or countries (i.e. disjoint groups). However, it is more intuitive to think of the clusters as the countries in this example. Econometric literature would suggest to cluster by groups where the 'treatment' occurs (Bertrand et. al, 2004). We assume that standard errors are correlated within countries but not between countries.
```{=latex}
\end{doublespace}
```
## GMM Application
```{=latex}
\begin{doublespace}
```
With an instrumental variables framework, we will use GMM to obtain estimations and confidence intervals for $\delta$. While the validity of the instrument has been discussed elsewhere, it is worth noting that with significant confidence, the relevance condition for the instrument has been met. By partialling out the fixed effects and covariates, we can obtain both a graphical and exact estimation of $\lambda$ (the first stage coefficient):

```{r relevance, message=F, warning=F, fig.width=10, fig.height=6, echo=F}
women.reg <- read_dta("Data/regression_matrix.dta")

#Partial out fixed effects and covariates
X.covariates = colnames(women.reg)[str_detect(colnames(women.reg), "lgdp|lpopulation")]
fixed.effects = colnames(women.reg)[str_detect(colnames(women.reg), "_I")]

women.reg.cleaned <- women.reg[complete.cases(women.reg[c(X.covariates,
                                       fixed.effects,
                                       "womenrep_lag", "z_lag")]),]

X.covariates = as.matrix(women.reg.cleaned[X.covariates])
fixed.effects = as.matrix(women.reg.cleaned[fixed.effects])

X.rel <- cbind(1, X.covariates, fixed.effects)

Y.rel <- women.reg.cleaned$womenrep_lag

#Partial out Y
Y.tilde.rel <- X.rel%*%(solve(t(X.rel)%*%X.rel)%*%t(X.rel)%*%Y.rel)-Y.rel

#Partial out Z
Z.rel <- women.reg.cleaned$z_lag
Z.tilde.rel <- X.rel%*%(solve(t(X.rel)%*%X.rel)%*%t(X.rel)%*%Z.rel)-Z.rel

#Partialled-out regression on Z
z.lm <- lm(Y.tilde.rel ~ 0+Z.tilde.rel)
z.beta = coef(z.lm)[1]
z.se = summary(z.lm)$coefficients[1,2]
z.ci = (z.beta + c(-1,1)*z.se*qnorm(0.975)) %>% round(5)

relevance.plot <- data.frame(x=Z.tilde.rel, y=Y.tilde.rel) %>%
  ggplot(aes(y=y, x=x))+
    geom_point(size=0.5)+
    geom_jitter(size=0.5, width=0.1, height=0.01)+
    geom_smooth(method = "lm", se=F, color="#44CCAA")+
    theme_minimal()+
    labs(
      x=expression(~tilde(Z)[ct-1]),
      y=expression(~tilde(W)[ct-1]),
      title = "Figure 3 - Relevance of Instrumental Variable Using Partialled-out Regressor and Regressand"
    )

relevance.plot
```
The regression coefficient is significantly positive:
`r z.beta %>% round(5)` (`r z.ci[1]`, `r z.ci[2]`), which, given that this identification is just-identified, implies $-\mathbb{E}[Z_{ct-1}\mathbb{X}_{ct}]$ has rank $k$, where $\mathbb{X_{ct}}$ is the full covariate X matrix (including fixed effects).

As for the moment condition, I assume,
```{=latex}
\end{doublespace}
```
$$
(7) \quad \mathbb{E}[m_i(\beta)]=\mathbb{E}[Z_{ct-1}(Y_{ct}-\mathbb{X}_{ct}'[\beta_0|\delta|\Omega_{1}|...|\Omega_{4}|\beta_{\text{Angola}}|...|\beta_{\text{Zimbabwe}}|\gamma_{1999}|...|\gamma_{2022}]')]=0
$$
$$
\iff\mathbb{E}[Z_{ct-1}v_{ct}|\mathbb{X}_{ct}]=0\quad\forall c,t
$$
```{=latex}
\begin{doublespace}
```
For identification, I also assume that the vector of parameters, $\beta$, uniquely solves (7).
```{=latex}
\end{doublespace}
```
## Estimation
```{=latex}
\begin{doublespace}
```
Assuming the conditions for identification and inference are met for the GMM identification strategy, we can use the standard two-stage least squares regression formula[^14]. Assuming homoskedasticity (using the formula for $\hat{V}$), we arrive at the following 95% C.I. for $\delta$, the local average treatment effect of interest:

[^14]: $(Z'X)^{-1}Z'Y$ for just-identified models. This was calculated using matrix algebra in R. All scripts for regression calculations and simulations can be found in <https://github.com/SamLeeBYU/Elections/CRVE>.

```{=latex}
\end{doublespace}
```
<!-- twosls(X, Y, Z) #These matrices are loaded in by setup.R -->
$$
-107.8425 \quad (-133.36956, -82.31544)
$$
```{=latex}
\begin{doublespace}
```
This means that (conditioned on the proportion of women being influenced by the instrument of years since suffrage was granted in the country), for every 1% increase in the average proportion of women in national legislatures of African and Arab countries, we would expect yearly per capita $CO_2$ emissions to decrease by 1.08 gigatons (-1.33, -0.82), on average[^15].

If we use the proposed cluster-robust estimators in Section III, we get significantly wider intervals (although still *statistically* significant):
```{=latex}
\end{doublespace}
```
```{r crve-intervals, fig.align='center', echo=F, warning=F, results='asis'}
cat(readLines("CRVE/crve-intervals.tex"), sep = "\n")
```
```{=latex}
\begin{doublespace}
```
The natural question that follows is whether or not we can trust the inference results on this regression output. Empirical results from Bertrand et. al (2004) suggest that the homoskedasticity assumption on $\epsilon_{ct}$ causes us to drastically understate the standard errors since $CO_2$ emissions are correlated within a country over time. Even if higher concentrations of women in national legislatures decrease carbon emissions through regulatory practice, there are reasonable arguments to suggest that the effects of those regulatory practices may not be fruit-bearing until years later down the road. After all, industries may need time to adjust their carbon-intensive practices into compliance. Thus, if the logic to cluster the standard errors by country prevails, which cluster-robust variance estimator should we use? Can we trust our inferences? I examine these questions in the next section.

[^15]: It is also worth noting that the only other covariates that were significant in the regression (given that standard errors were calculated assuming homoskedasticity) besides fixed effects were the coefficients on $\text{GDP}_{ct-2}$ and $\text{Population}_{ct-2}$, which were both significantly positive, though not nearly as high in magnitude as $\hat{\delta}$.
```{=latex}
\end{doublespace}
```
# V. Monte Carlo Analysis to Assess Performance of Cluster-Robust Variance Estimators
```{=latex}
\begin{doublespace}
```
To assess which (if any) of my proposed cluster-robust variance estimators are appropriate for a (just-identified) instrumental variables framework in a GMM setting, I will apply a Monte Carlo analysis to our empirical example. The goal of this analysis will be to see if (i) the results from our empirical analysis can be reasonably trusted. Assuming the conditions for identification hold, can we accurately assess the (local average treatment) effect that electing more women to national legislatures in African and Arab countries has on yearly per capita $CO_2$ emissions? This analysis also serves to (ii) analyze the long-run performance of the cluster-robust variance estimators I proposed in Section III.
```{=latex}
\end{doublespace}
```
## Data-Generating Process
```{=latex}
\begin{doublespace}
```
The data-generating process will proceed as follows. While it is possible to choose a pre-determined $\beta_0$ vector as other authors have previously done in the literature and generate completely new data according to a similar framework outlined by equations (3) and (4), I wish to simulate data using the same distribution of my sample data in my empirical example. To do this, I will perform a bootstrap sampling of $Y_{ct}$. By leaving my covariates, instrument, and treatment data the same as in the sample data, I will replace the empirical sample $Y_{ct}$ with a new bootstrap sample of $Y_{ct}$, which I will define as $\overset{\circ}{Y}_{ct}$. By running GMM estimation for two-stage least squares for each of the proposed CRVEs, I will assess the long-run empirical rejection rates for each of my estimators using many iterations (simulations) of $\overset{\circ}{Y}_{ct}$. Mathematically[^16], I will define $\overset{\circ}{Y}_{ct}$ as,

[^16]: Note, with the probability statement defined as it is, this means that each element in ${Y}_{ct}$ is sampled with a uniform probability mass (or, equivalently, sampled with replacement with equal probability) to create $\overset{\circ}{Y}_{ct}$.
```{=latex}
\end{doublespace}
```
$$
\overset{\circ}{Y}_{ct}=\left\{\overset{\circ}{y_{ct}}\right\}_{c=\text{Angola, } t=1999}^{c=\text{Zimbabwe, t=2022}}=\left\{ \overset{\circ}{y_{\text{Angola, } 1999}}, \overset{\circ}{y_{\text{Angola, } 2000}},...,\overset{\circ}{y_{\text{Zimbabwe, } 2022}} \right\}
$$
$$
Pr\left( \overset{\circ}{y}_{ct} = y_{ct} \right) = \frac{1}{n}, \quad \forall c,t
$$
```{r dgp-vis, echo=F, fig.width=16, fig.height=8}
original.Y <- data.frame(x=Y) %>%
  ggplot(mapping=aes(x=x))+
  geom_histogram(binwidth = 1, fill = "#44CCAA", color = "black") +
    labs(title = expression(Y[ct]), x = "Yearly (Gigatons of) per Capita Carbon Emissions", y = "Frequency") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          axis.text = element_text(size = 10),
          axis.title = element_text(size = 12),   
          panel.grid.minor = element_blank(),    
          panel.grid.major = element_line(color = "lightgrey"))

boostrap.Y <- data.frame(x=DGP(clusters=clusters, Y, cluster=F)) %>%
  ggplot(mapping=aes(x=x))+
  geom_histogram(binwidth = 1, fill = "#44CCAA", color = "black") +
    labs(title = expression("Bootstrap Sample of "~Y[ct]), x = "Yearly (Gigatons of) per Capita Carbon Emissions", y = "Frequency",
         caption = "Figure 4 - Comparison between original dependent variable and bootstrapped dependent variable.") +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          axis.text = element_text(size = 10),
          axis.title = element_text(size = 12),   
          panel.grid.minor = element_blank(),    
          panel.grid.major = element_line(color = "lightgrey"))

#ggsave("Graphs/DGP.png", original.Y + boostrap.Y, width=16, height=8)

original.Y + boostrap.Y
```
## Simulation
```{=latex}
\begin{doublespace}
```
The DGP simulates placebo vectors for $Y_{ct}$. If the results from the main regression were not obtained by random chance, we would expect that $\overset{\circ}{Y}_{ct}$ would testify to this effect. Thus, on average---assuming the standard errors are sized correctly---using a 95% confidence interval, we would expect the empirical rejection rate of estimators for $\hat{\delta}$ to approach 5% using $\overset{\circ}{Y}_{ct}$ in place of $Y_{ct}$. However, if the results were obtained by some misspecification or some other unobserved factor, it may be that the empirical rejection rate may not converge to 5%, and the GMM variance estimators may fail to deliver on their promise. In such a scenario, we will not be able to put much confidence in the two-stage difference-in-differences regression results. Simultaneously, if we can put our confidence in the regression results, we can also evaluate the empirical performance of the proposed cluster-robust variance estimators and determine which estimator may be preferential in a just-identified GMM setting that uses instrumental variables.

I run 10,000 simulations, generating a new $\overset{\circ}{Y}_{ct}$ for each simulation. I then calculate the estimates for $\delta$ (the local average treatment effect of interest) and the respective 95% confidence intervals using the three proposed CRVEs in Section III as well as the 95% confidence interval using the equation for $\hat{V}$, which assumes a homoskedastic variance on $\epsilon_{ct}$, as a baseline measurement. Notably, calculating $\overset{iv}{CV}_{2}$ and $\overset{iv}{CV}_{3}$ requires the calculation of $\overset{iv}{M}_{gg}$. In order to calculate this matrix, it is requisite to partial out fixed effects to avoid singularity (MacKinnon et. al, 2022).
```{=latex}
\end{doublespace}
```
## Simulation Results
```{r simulation, echo=F, fig.width=14, fig.height=8}
load("CRVE/sims.RData")

sims.delta <- data.frame(x=sapply(sims[,1]+sims[,2], median)) %>%
  ggplot(mapping=aes(x=x))+
  geom_histogram(binwidth = 5, fill = "#44CCAA", color = "black") +
    labs(title = expression("Distribution of "~hat(delta)), x = expression(hat(delta)), y = "Frequency",
         caption = "Figure 5 - Sample distribution of the simulated estimators for the local average treatment effect") +
  geom_vline(xintercept = 0, linetype=2)+
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
          axis.text = element_text(size = 10),
          axis.title = element_text(size = 12),   
          panel.grid.minor = element_blank(),    
          panel.grid.major = element_line(color = "lightgrey"))

#ggsave("Graphs/SimulatedDelta.png", sims.delta, width=14, height=8)

sims.delta
```
```{r sim-results, fig.align='center', echo=F, warning=F, results='asis'}
cat(readLines("CRVE/sim-results.tex"), sep = "\n")
```
# VI. Implications and Conclusion
```{=latex}
\begin{doublespace}
```
The results from the Monte Carlo simulation imply that using $\overset{iv}{CV}_{1}$ may be the best cluster-robust variance estimator in just-identified instrumental variables settings. The $\overset{iv}{CV}_{1}$ estimator uses a degrees of freedom correction. This implies that since the empirical rejection rate seems to hover slightly above 5%, the cluster robust variance formula employed by Stata may systematically overestimate the standard errors when the option `cluster` is specified on an `ivregress` command[^19]. $\overset{iv}{CV}_{2}$ and $\overset{iv}{CV}_{3}$ seem to routinely under reject. However, it seems that MacKinnon et. al are also aware of this unique feature of the jackknife estimator. The authors still favor this estimator over the first two alternatives for other reasons, including computational motives. This paper reaffirms this characteristic about $\overset{iv}{CV}_{3}$. It may also be possible that the characteristics seen in the analysis are unique to the data-generating process used or the data itself. More research and analysis may prove beneficial on this front. Additionally, this simulation was performed with $G=64$ (64 unique countries in the analytic sample). It may be possible that the different estimators perform differently under different conditions. Furthermore, when clustering is impractical either via theoretical analysis or empirically infeasible, this study shows that the two-stage least squares estimator (given that assumptions for identification are properly met) seems to work just as fine. Overall, while this paper cannot hope to be the final word of theoretical and empirical work on cluster-robust variance estimators, I do hope this paper has served to enforce the credibility of inference employed by influential papers that cluster their standard errors.

Finally, this analysis adds credibility to the empirical findings of the applied example presented in this paper. The results of the Monte Carlo analysis imply that the test used to evaluate whether $\hat{\delta}$ is significant seems appropriately sized. Given that electing higher proportions of women in national legislatures in African and Arab countries truly has no effect on yearly per capita $CO_2$ emissions, using $\overset{iv}{CV}_{1}$[^18], there is 0.021% chance (p-value < 0.001) that we would observe an effect as large or larger by random. Given these results, we can (confidently) reaffirm the effect found by Mavisakalyan and Tarverdi. While the exact mechanism through which women's
presence influences climate policy remains open for further
investigation, these findings hold significant policy implications. It seems clear when more women enter
the national legislature, we see lower per capita $CO_2$ emissions.

[^18]: The 95% confidence interval using $\overset{iv}{CV}_{1}$ yields (-164.8728, -50.8122), or, on the 1% scale, (-1.65, -5.08).
[^19]: Curiously, the authors of MacKinnon et. al use a critical value of $t_{1-\alpha/2, df=G-1}$, whereas Stata assumes a Normal critical value. Simulation results were calculated using Normal critical values. Using a critical value adjusted for $G$ such as suggested by MacKinnon et. al yields slightly lower rejection rates in the simulation.

```{=latex}
\end{doublespace}
```
\newpage
# References

Bertrand, M., Duflo, E., & Mullainathan, S. (2004). How much should we
trust differences-in-differences estimates?
<https://www.jstor.org/stable/25098683>

Bell, R.M.m McCaffrey, D.F., (2002). Bias reduction in standard errors for linear regression with multi-stage samples. Surv. Methodol. 28, 169-181.

Emissions Database for Global Atmospheric Research (EDGAR). (n.d.).
Edgar online. Retrieved from <https://edgar.jrc.ec.europa.eu/>

European Commission, Joint Research Centre (JRC). (2023). Emissions gap
report 2023. Publications Office of the European Union.
doi:10.2760/481442 <https://edgar.jrc.ec.europa.eu/report_2023>

Inter-Parliamentary Union (IPU). (n.d.). Women in national parliaments.
\[Data set\]. Retrieved from <https://data.ipu.org/womon-ranking>

MacKinnon, J. G., Nielsen, M. ., & Webb, M. D. (2023). Cluster-robust inference: A guide to empirical practice. Journal of Econometrics, 47(4), 4144561

Mavisakalyan, A., & Tarverdi, Y. (2019). Gender and climate change: Do
female parliamentarians make a difference? Global Labor Organization
(GLO) Discussion Paper No. 221.
<https://www.sciencedirect.com/science/article/pii/S0176268017304500>

Medlicott, E. (2021, March 8). The five devastating reasons climate
change affects women more than men. Retrieved from
<https://www.euronews.com/green/2021/11/09/the-five-devastating-reasons-climate-change-affects-women-more-than-men>

Mirziyoyeva, N., & Salahodjaev, D. (2023). Does representation of women
in parliament promote economic growth? Considering evidence from Europe
and Central Asia. Frontiers in Political Science, 4(2). \[DOI:
10.3389/fpos.2023.1203221\]
<https://www.frontiersin.org/articles/10.3389/fpos.2023.1120287/full>

Schaeffer, Katherine (2020, October 5). Pew Research Center. Key facts
about women's suffrage around the world, a century after U.S. ratified
19th Amendment. Retrieved April 1, 2024, from
<https://www.pewresearch.org/short-reads/2020/10/05/key-facts-about-womens-suffrage-around-the-world-a-century-after-u-s-ratified-19th-amendment/>

Sellers, S. (2016). Gender and climate change: A closer look at existing
evidence. Global Gender and Climate Alliance (GGCA).
<https://wedo.org/gender-and-climate-change-a-closer-look-at-existing-evidence-ggca/>

Skaaning, Svend-Erik; John Gerring; Henrikas Bartusevicius, 2015, "A
Lexicial Index of Electoral Democracy",
<https://doi.org/10.7910/DVN/29106>, Harvard Dataverse, V6

United Nations Development Programme (UNDP). (2022, March 8).
International Women's Day 2022 "Gender equality today for a sustainable
tomorrow". Retrieved from
<https://www.undp.org/speeches/undp-administrators-statement-international-womens-day-2023-8-march>

U.S. Department of State. (n.d.). Countries and Areas List. Retrieved
from <https://www.state.gov/countries-and-areas-list/>

The World Bank. (n.d.). World Bank Open Data. Retrieved from
<https://databank.worldbank.org>